---
title: "SARP123+CSGA LAAA Results - Strat by Sex"
author: "SDS"
date: "Last compiled on `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: kable
    code_folding: hide
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = paste0(Sys.getenv("RKJCOLLAB"), "/Collabs/ortega/reports")) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(
  root.dir = paste0(Sys.getenv("RKJCOLLAB"), "/Collabs/ortega"))

# Common packages
library(tidyverse)
library(knitr)
library(biomaRt)
library(DT)
library(conflicted)
library(scales)
library(patchwork)

#  For locuszoomr
library(locuszoomr)
library(ensembldb)
library(AnnotationHub)
library(GGally)
library(RColorBrewer)
library(ggpubr)
library(IRanges)
library(snpStats)

# Set display options
options(digits = 4)  # only 2 digits after decimal

```

# **Overview**

Makes combined locus zoom and phenotype plots from LAAA/fine mapping results.


**LAAA model formula:**

pheno ~ group + covariates + RFMIX_GW_AFR + allele_dose + afr_dose + allele_afr_dose

For locus zoom plots, LD is calculated within our study population using
R package snpStats.

Significance threshold can be changed, but current script follows Dr. Ortega's
and Michelle's previous paper: P < 10^-3 (0.001). Significant SNPs are those
whose allele dose p-value and/or interaction dose p-value meets this threshold.

# **Load Data**

```{r load_data, message = F}

# TODO: update all paths in this section

result_dir <- "data/laaa_wgs/strat_by_sex"
result_list <- list.files(
  path = result_dir,
  pattern = ".*_assoc_chr[0-9]+\\.txt",
  recursive = T)

# Extract varying values
sex <- gsub("output_", "", dirname(dirname(result_list)))
pheno_chr <- str_match(
  basename(result_list), "^(.*)_assoc_chr([0-9]{1,2})\\.txt$")
pheno <- pheno_chr[,2]
chr <- pheno_chr[,3]

# Read in data
invisible(map2(result_list, paste0(pheno, "_chr", chr, "_", sex), function(file, name) {
  data <- read.table(paste0(result_dir, "/", file), header = TRUE)
  assign(name, data, envir = .GlobalEnv)
}))

# Function to read in dose frames
read_dose_frames <- function(dir_name) {
    df_list <- list.files(dir_name, full.names = T)
    
    df_all <- Reduce(rbind, lapply(df_list, function(file) {
      file_base <- basename(file)
      df_name <- gsub(
        ".+[[:digit:]]+_", "",
        gsub("\\.txt", "", file_base))
      df <- read_delim(file) %>%
        dplyr::mutate(type = df_name) %>%
        relocate(type)
      return(df)
  }))
  return(df_all)
}

# Add paths to each directory that contains dose frames
dose_dir <- "data/laaa_wgs/strat_by_sex/dose_frames"
dose_list <- list.dirs(
  path = dose_dir,
  recursive = T)
dose_list <- grep("hg38", dose_list, value = T)
chr_list <- gsub(".*_chr", "", dose_list)

# Read in data
invisible(map2(dose_list, chr_list, function(dose, chr) {
  data <- read_dose_frames(dose)
  assign(paste0("df_chr", chr), data, envir = .GlobalEnv)
}))

# Add paths to same phenotype files used while running LAAA
# Using the male/female combined here, but ran LAAA with split files
pheno_mff_female <- read_delim(
  "data/pheno/strat_by_sex/SARP123_CSGA_234_laaa_maxFEV1_FVC_pheno_female.txt")
pheno_mff_male <- read_delim(
  "data/pheno/strat_by_sex/SARP123_CSGA_114_laaa_maxFEV1_FVC_pheno_male.txt")
pheno_mfvc_female <- read_delim(
  "data/pheno/strat_by_sex/SARP123_CSGA_234_laaa_maxFVC_pheno_female.txt")
pheno_mfvc_male <- read_delim(
  "data/pheno/strat_by_sex/SARP123_CSGA_114_laaa_maxFVC_pheno_male.txt")

# Update path to script that sets main paper colors
source("/Users/slacksa/repos/local_ancestry_analysis/laaa_fine_mapping/set_main_paper_colors.R")

# Set significance thresholds - both normal and nominal
sig_thresh <- 0.001
nom_sig_thresh <- 0.01

```

Get plotting coordinates:

```{r, message=F}
# Get plotting coordinates to include combined & sex-stratified analyses
df_chr8_comb <- read_delim(
  "data/laaa_wgs/dose_frames/region_hg38_maxFVC_chr8/maxFVC_chr_8_afr_dose.txt")

reg_chr8_min <- min(df_chr8$position, df_chr8_comb$position)
reg_chr8_max <- max(df_chr8$position, df_chr8_comb$position)
y_max <- 1e-7
```

Get range of each term p-values:

```{r}
# Get range of each term p-values to compare with combined analysis
# Using chr8 maxFVC peak for this
p_cols <- c("allele_dose_p", "afr_dose_p", "allele_afr_dose_p", "anova_p")
maxFVC_chr8_female_max <- maxFVC_chr8_female %>%
  dplyr::summarize(across(all_of(p_cols),
      list(min = ~min(.x, na.rm = TRUE), max = ~max(.x, na.rm = TRUE))))
maxFVC_chr8_female_max
maxFVC_chr8_male_max <- maxFVC_chr8_male %>%
  dplyr::summarize(across(all_of(p_cols),
      list(min = ~min(.x, na.rm = TRUE), max = ~max(.x, na.rm = TRUE))))

write_tsv(
  maxFVC_chr8_female_max,
  "data/laaa_wgs/strat_by_sex/output_female/min_max_pvals_maxFVC_chr8.txt")
write_tsv(
  maxFVC_chr8_male_max,
  "data/laaa_wgs/strat_by_sex/output_male/min_max_pvals_maxFVC_chr8.txt")

```


# **Functions**

```{r}
# Get significant results
# Can run to get significant for given p-value
get_sig <- function(df, thresh, pval) {
  # Get significant results
  df_sig <- df %>% dplyr::filter(!!sym(pval) < thresh)
  
  # Write abs(max) column name in new column
  df_sig$max_abs_beta <- apply(
    df_sig[, c("allele_dose_beta", "afr_dose_beta", "allele_afr_dose_beta")],
    1, function(row) {
      cols <- c("allele_dose_beta", "afr_dose_beta", "allele_afr_dose_beta")
      cols[which.max(abs(row))]
  })
  return(df_sig)
}

# Make plotting function - all p-values in one plot
plot_sig <- function(win_df, min_pos = NULL, max_pos = NULL, y_max = 1e-7) {

  win_df_lg <- win_df %>%
    dplyr::select(position, allele_dose_p, afr_dose_p, allele_afr_dose_p, anova_p) %>%
    pivot_longer(cols = c(allele_dose_p, afr_dose_p, allele_afr_dose_p, anova_p),
                 names_to = "type",
                 values_to = "pval")
  
  win_df_sig <- subset(win_df_lg, pval < sig_thresh & type == "anova_p")
  win_df_nom_sig <- subset(win_df_lg, pval < nom_sig_thresh & type == "anova_p")
  
  if (!is.null(min_pos) & !is.null(max_pos)) {
    ggplot(win_df_lg, aes(position, -log10(pval))) +
      geom_point() +
      geom_point(data = subset(win_df_lg, position %in% win_df_sig$position),
                 aes(color = "orange")) +
      labs(x = "Position", y = "-log10(P-Value)", color = paste0("Anova P < ", sig_thresh)) +
      facet_wrap(~ type) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_x_continuous(limits = c(min_pos, max_pos)) +
      scale_y_continuous(limits = c(NA, -log10(y_max)))
  } else {
    ggplot(win_df_lg, aes(position, -log10(pval))) +
      geom_point() +
      geom_point(data = subset(win_df_lg, position %in% win_df_sig$position),
                 aes(color = "orange")) +
      labs(x = "Position", y = "-log10(P-Value)", color = paste0("Anova P < ", sig_thresh)) +
      facet_wrap(~ type) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_y_continuous(limits = c(NA, -log10(y_max)))
  }
}

# Plot split phenotype
# For testing
# pheno_data <- pheno_mfvc
# pheno_data_id_col <- "TopMed_ID"
# dose_frame <- df_chr8
# sig_data <- win_chr8_a_sig
# position <- win_chr8_a_sig_peaks$position[1]
# pheno <- "maxFVC"
# colors = c(color_lfvc, color_bfvc, color_maxfvc)
fun_split <- function(x, data, pheno) {
  return(data.frame(
    y = min(data[[pheno]]) - (0.1*(max(data[[pheno]]) - min(data[[pheno]]))),
    label = length(x)))  # x comes from group = Asthma
}
plot_split <- function(
    pheno_data, pheno_data_id_col, dose_frame, sig_data, position, pheno, colors) {
 
  # Merge pheno and dose frame
  df_afr <- dose_frame %>% dplyr::filter(type == "afr_dose") %>%
    dplyr::select(-type, -ref, -alt) %>%
    column_to_rownames(var = "position")
  df_afr_t <- as.data.frame(t(df_afr))
  colnames(df_afr_t) <- paste0(colnames(df_afr_t), "_afr")
  df_afr_t <- rownames_to_column(df_afr_t, var = "ID")
  
  df_allele <- dose_frame %>% dplyr::filter(type == "allele_dose") %>%
    dplyr::select(-type, -ref, -alt) %>%
    column_to_rownames(var = "position")
  df_allele_t <- as.data.frame(t(df_allele))
  colnames(df_allele_t) <- paste0(colnames(df_allele_t), "_allele")
  df_allele_t <- rownames_to_column(df_allele_t, var = "ID")
  
  # Convert ID columns to chr before joining
  pheno_data[[pheno_data_id_col]] = as.character(pheno_data[[pheno_data_id_col]])
  df_afr_t$ID <- as.character(df_afr_t$ID)
  df_allele_t$ID <- as.character(df_allele_t$ID)
  
  pheno_df_temp <- inner_join(
    pheno_data,
    df_afr_t,
    by = setNames("ID", pheno_data_id_col))  # setNames flips order
  pheno_df <- inner_join(
    pheno_df_temp,
    df_allele_t,
    by = setNames("ID", pheno_data_id_col))  # setNames flips order
  
  # Get additional details for plot from sig_data
  driving_term <- sig_data$max_abs_beta[sig_data$position == position]
  
  # Make format of allele dose that have alt & ref alleles
  alt <- sig_data$alt[sig_data$position == position]
  ref <- sig_data$ref[sig_data$position == position]
  allele_labels <- c(
    `0` = paste0(ref, ref),
    `1` = paste0(ref, alt),
    `2` = paste0(alt, alt))
  
  plot <- ggplot(pheno_df, aes(
    x = factor(!!rlang::sym(paste0(position, "_allele"))),
    y = !!rlang::sym(pheno))) +
  geom_boxplot(aes(fill = "Combined"),
               varwidth = TRUE, alpha = 0.7, outlier.shape = 21) +
  geom_boxplot(aes(fill = factor(!!rlang::sym(paste0(position, "_afr")))),
               varwidth = TRUE, alpha = 0.7, outlier.shape = 21) +
  labs(x = "Genotype",
       y = paste0(
         gsub("_", "/", gsub("m", "M", pheno)),
         ifelse(str_detect(pheno, "_"), " Ratio", " (mL)")),
       fill = "Copies Local African Ancestry",
       title = paste0(
         "chr", sig_data$chr, ":", position, ":", ref, ":", alt, ", Driving Term: ", driving_term)) +
  scale_x_discrete(labels = allele_labels) +
  scale_y_continuous(labels = label_comma()) +
  scale_fill_manual(values = c("Combined" = "transparent",  # Transparent box plot
                             "0" = colors[1], "1" = colors[2], "2" = colors[3]),
                    labels = c("Combined" = "Combined", "0" = "0",
                               "1" = "1", "2" = "2")) +
  theme_bw() +
  theme(legend.position = "bottom") +
  stat_summary(fun.data = fun_split, fun.args = list(data = pheno_df, pheno = pheno),
               geom = "text", aes(group = !!rlang::sym(paste0(position, "_afr"))),
               position = position_dodge(0.5), hjust = 0.5)
  return(plot)
}

# Get LD
# peak_snp <- win_chr5_a_sig_peaks$position[1]
# win_sig <- win_chr5_a_sig
# dose_frame <- df_chr5
# pval <- "allele_dose_p"
calc_peak_ld <- function(peak_snp, win_sig, dose_frame, pval) {
  # Get peak snp
  peak <- paste0(
    win_sig$position[which.min(win_sig[[pval]])], "_allele")
  
  # Modify dose frame, use to create SnpMatrix
  df_allele_t <- as.data.frame(t(dose_frame %>%
    dplyr::filter(type == "allele_dose") %>%
    dplyr::select(-type, -ref, -alt) %>%
    column_to_rownames(var = "position")))
  snp_m <- new("SnpMatrix", as.matrix(df_allele_t))
  
  # Get pairwise LD with peak SNP
  ld <- ld(snp_m[, which(colnames(snp_m) == peak_snp), drop = F],
           snp_m, stats = c("R.squared"))
  
  # LD R2 column name is just position of peak SNP calculated with respect to
  ld_t <- as.data.frame(t(ld)) %>%
    rownames_to_column("pos") %>%
    dplyr::mutate(pos = as.integer(gsub("_allele", "", pos)))
  
  return(ld_t)
}

# Plot locus zoom
# win_ld <- win_chr8_int_ld
# chr <- "8"
# peak_snp <- chr8_int_peak
# pval <- "allele_afr_dose_p"
# win_ld <- win_chr5_a_ld
# chr <- "5"
# peak_snp <- chr5_a_peak
# pval <- "allele_dose_p"
plot_locuszoom <- function(win_ld, chr, peak_snp, pval) {
  # Remove NA pvalues
  # TODO: revisit and confirm this as expected!
  win_ld <- win_ld[!is.na(win_ld[[pval]]), ]
  # Plot
  loc <- locus(data = win_ld, chrom = "chr", p = pval, pos = "position",
             ens_db = ensDb_v111, seqname = chr, labs = "position",
             xrange = c(min(win_ld$position), max(win_ld$position)),
             LD = as.character(peak_snp))
  loc_plot <- locus_ggplot(loc, filter_gene_biotype = "protein_coding",
             pcutoff = sig_thresh, legend_pos = "top")
  return(loc_plot)
}

```

```{r}
# win_df_female <- maxFVC_chr5_female
# win_df_male <- maxFVC_chr5_male
# pval <- "allele_dose_p"
# test <- plot_miami(maxFVC_chr5_female, maxFVC_chr5_male, "allele_dose_p")
plot_miami <- function(win_df_top, win_df_bottom, pval,
                       min_pos = NULL, max_pos = NULL, y_max = 1e-7) {
  # Prepare top (female) data
  top_df <- win_df_top %>%
    dplyr::select(position, pval = all_of(pval)) %>%
    dplyr::mutate(logp = log10(pval),
           highlight = pval < sig_thresh)
  
  # Prepare bottom (male) data
  bottom_df <- win_df_bottom %>%
    dplyr::select(position, pval = all_of(pval)) %>%
    dplyr::mutate(logp = log10(pval),
           highlight = pval < sig_thresh)

  ## Top plot (Female)
  p_top <- ggplot(top_df, aes(x = position, y = -logp)) +
    geom_point(aes(color = highlight), alpha = 0.6) +
    scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black"), guide = "none") +
    geom_hline(yintercept = -log10(sig_thresh), linetype = "dashed", color = "blue") +
    scale_y_continuous(limits = c(0, -log10(y_max))) +
    labs(y = bquote("Female " ~ -log[10](.(pval))), x = "Position") +
    theme_minimal() +
    theme(axis.text.x = element_blank(),
          plot.title = element_text(hjust = 0.5))

  ## Bottom plot (Male), flipped with scale_y_reverse()
  p_bottom <- ggplot(bottom_df, aes(x = position, y = -logp)) +
    geom_point(aes(color = highlight), alpha = 0.6) +
    scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black"), guide = "none") +
    geom_hline(yintercept = -log10(sig_thresh), linetype = "dashed", color = "blue") +
    scale_y_reverse(limits = c(-log10(y_max), 0)) +
    labs(y = bquote("Male " ~ -log[10](.(pval))), x = "Position") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))

  ## Stack with patchwork
  p <- p_top / p_bottom + plot_layout(heights = c(1, 1)) & 
  theme(plot.margin = margin(0, 0, 0, 0))

  # Set x limits if provided
  if (!is.null(min_pos) & !is.null(max_pos)) {
    p <- p + scale_x_continuous(limits = c(min_pos, max_pos))
  }
  
  return(p)
}


```


# **Results**

For each window, first check for overall model significance. For each variant
that is significant, then identify the effect with the largest absolute value
of the test statistics to determine which part of the LAAA model is driving
that significance.

## **MaxFVC Chr5**

### **All Model P-Values**

```{r}
plot_sig(maxFVC_chr5_female)
plot_sig(maxFVC_chr5_male)
```

### **Miami Plots**

```{r out.width = '50%', fig.show = 'hold'}
plot_miami(maxFVC_chr5_female, maxFVC_chr5_male, "allele_dose_p")
plot_miami(maxFVC_chr5_female, maxFVC_chr5_male, "allele_afr_dose_p")
```


### **Allele & Interaction Results**

#### Female

##### **Allele Plot with Phenotype**

```{r}
# SET FOR WHOLE SECTION
pheno <- "maxFVC"
chr <- "5"
sex <- "female"

# Get sig LAAA results & format
prefix <- paste0(pheno, "_chr", chr, "_", sex)
invisible(assign(
  paste0(prefix, "_a_sig"),
  get_sig(get(prefix), sig_thresh, "allele_dose_p")) %>%
  dplyr::mutate(chr = chr) %>%
  dplyr::arrange(allele_dose_p))
invisible(assign(
  paste0(prefix, "_int_sig"),
  get_sig(get(prefix), sig_thresh, "allele_afr_dose_p")) %>%
  dplyr::mutate(chr = chr) %>%
  dplyr::arrange(allele_afr_dose_p))

get(paste0(prefix, "_a_sig")) %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)

```

One for each SNP crossing significance threshold, arranged from most to least
significant. Only showing the top 6.

```{r out.width = '50%', fig.show = 'hold'}
# Inputs to plot_split should be
  # pheno_mfvc = pheno dataframe prepared above
  # "TopMed_ID" = name of ID column in pheno file
  # df_chr17 = dose frame dataframe prepared above
  # win_chr17_sig = significant LAAA results dataframe prepared above
  # .x = significant positions from win_chr17_sig prepared above
  # "maxFVC" = name of phenotype plotting - may need to update the plotting
    # function/units if changing this
  # colors = three plotting colors for manuscript
colors = c(color_lfvc, color_bfvc, color_maxfvc)


walk(get(paste0(prefix, "_a_sig"))$position, ~ {
  print(plot_split(
    get(paste0("pheno_mfvc_", sex)),
    "TopMed_ID",
    get(paste0("df_chr", chr)),
    get(paste0(prefix, "_a_sig")), .x, "maxFVC", colors))
})

```

##### **Interaction Plot with Phenotype**

```{r}
get(paste0(prefix, "_int_sig")) %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)
```


```{r out.width = '50%', fig.show = 'hold'}
walk(get(paste0(prefix, "_int_sig"))$position, ~ {
  print(plot_split(
    get(paste0("pheno_mfvc_", sex)),
    "TopMed_ID",
    get(paste0("df_chr", chr)),
    get(paste0(prefix, "_int_sig")), .x, "maxFVC", colors))
})

```

#### Male

##### **Allele Plot with Phenotype**

```{r}
# SET FOR WHOLE SECTION
pheno <- "maxFVC"
chr <- "5"
sex <- "male"

# Get sig LAAA results & format
prefix <- paste0(pheno, "_chr", chr, "_", sex)
invisible(assign(
  paste0(prefix, "_a_sig"),
  get_sig(get(prefix), sig_thresh, "allele_dose_p")) %>%
  dplyr::mutate(chr = chr) %>%
  dplyr::arrange(allele_dose_p))
invisible(assign(
  paste0(prefix, "_int_sig"),
  get_sig(get(prefix), sig_thresh, "allele_afr_dose_p")) %>%
  dplyr::mutate(chr = chr) %>%
  dplyr::arrange(allele_afr_dose_p))

get(paste0(prefix, "_a_sig")) %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)

```

One for each SNP crossing significance threshold, arranged from most to least
significant. Only showing the top 6.

```{r out.width = '50%', fig.show = 'hold'}
# Inputs to plot_split should be
  # pheno_mfvc = pheno dataframe prepared above
  # "TopMed_ID" = name of ID column in pheno file
  # df_chr17 = dose frame dataframe prepared above
  # win_chr17_sig = significant LAAA results dataframe prepared above
  # .x = significant positions from win_chr17_sig prepared above
  # "maxFVC" = name of phenotype plotting - may need to update the plotting
    # function/units if changing this
  # colors = three plotting colors for manuscript
colors = c(color_lfvc, color_bfvc, color_maxfvc)

walk(get(paste0(prefix, "_a_sig"))$position, ~ {
  print(plot_split(
    get(paste0("pheno_mfvc_", sex)),
    "TopMed_ID",
    get(paste0("df_chr", chr)),
    get(paste0(prefix, "_a_sig")), .x, "maxFVC", colors))
})

```

##### **Interaction Plot with Phenotype**

```{r}
get(paste0(prefix, "_int_sig")) %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)
```


```{r out.width = '50%', fig.show = 'hold'}
walk(get(paste0(prefix, "_int_sig"))$position, ~ {
  print(plot_split(
    get(paste0("pheno_mfvc_", sex)),
    "TopMed_ID",
    get(paste0("df_chr", chr)),
    get(paste0(prefix, "_int_sig")), .x, "maxFVC", colors))
})

```


## **MaxFVC Chr8**

### **All Model P-Values**

```{r}
plot_sig(maxFVC_chr8_female, reg_chr8_min, reg_chr8_max)
plot_sig(maxFVC_chr8_male, reg_chr8_min, reg_chr8_max)
```

### **Miami Plots**

```{r out.width = '50%', fig.show = 'hold'}
plot_miami(maxFVC_chr8_female, maxFVC_chr8_male, "allele_dose_p")
plot_miami(maxFVC_chr8_female, maxFVC_chr8_male, "allele_afr_dose_p")
```

### **Allele & Interaction Results**

*TODO: make this a function & update all other sections!*

#### Female

##### **Allele Plot with Phenotype**

```{r}
# SET FOR WHOLE SECTION
pheno <- "maxFVC"
chr <- "8"
sex <- "female"

# Get sig LAAA results & format
prefix <- paste0(pheno, "_chr", chr, "_", sex)
invisible(assign(
  paste0(prefix, "_a_sig"),
  get_sig(get(prefix), sig_thresh, "allele_dose_p")) %>%
  dplyr::mutate(chr = chr))
invisible(assign(
  paste0(prefix, "_int_sig"),
  get_sig(get(prefix), sig_thresh, "allele_afr_dose_p")) %>%
    dplyr::mutate(chr = chr))

get(paste0(prefix, "_a_sig")) %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p) %>%
  dplyr::arrange(allele_dose_p)

a_sig_pos <- get(paste0(prefix, "_a_sig")) %>%
  dplyr::arrange(allele_dose_p) %>%
  dplyr::pull(position)

```

One for each SNP crossing significance threshold, arranged from most to least
significant. Only showing the top 6.

```{r out.width = '50%', fig.show = 'hold'}
# Inputs to plot_split should be
  # pheno_mfvc = pheno dataframe prepared above
  # "TopMed_ID" = name of ID column in pheno file
  # df_chr17 = dose frame dataframe prepared above
  # win_chr17_sig = significant LAAA results dataframe prepared above
  # .x = significant positions from win_chr17_sig prepared above
  # "maxFVC" = name of phenotype plotting - may need to update the plotting
    # function/units if changing this
  # colors = three plotting colors for manuscript
colors = c(color_lfvc, color_bfvc, color_maxfvc)

walk(a_sig_pos[1:6], ~ {
  print(plot_split(
    get(paste0("pheno_mfvc_", sex)),
    "TopMed_ID",
    get(paste0("df_chr", chr)),
    get(paste0(prefix, "_a_sig")), .x, "maxFVC", colors))
})

```

##### **Interaction Plot with Phenotype**

```{r}
get(paste0(prefix, "_int_sig")) %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p) %>%
  dplyr::arrange(allele_afr_dose_p)

int_sig_pos <- get(paste0(prefix, "_int_sig")) %>%
  dplyr::arrange(allele_afr_dose_p) %>%
  dplyr::pull(position)
```


```{r out.width = '50%', fig.show = 'hold'}
walk(int_sig_pos, ~ {
  print(plot_split(
    get(paste0("pheno_mfvc_", sex)),
    "TopMed_ID",
    get(paste0("df_chr", chr)),
    get(paste0(prefix, "_int_sig")), .x, "maxFVC", colors))
})

```




#### Male

##### **Allele Plot with Phenotype**

```{r}
# SET FOR WHOLE SECTION
pheno <- "maxFVC"
chr <- "8"
sex <- "male"

# Get sig LAAA results & format
prefix <- paste0(pheno, "_chr", chr, "_", sex)
invisible(assign(
  paste0(prefix, "_a_sig"),
  get_sig(get(prefix), sig_thresh, "allele_dose_p")) %>%
  dplyr::mutate(chr = chr))
invisible(assign(
  paste0(prefix, "_int_sig"),
  get_sig(get(prefix), sig_thresh, "allele_afr_dose_p")) %>%
  dplyr::mutate(chr = chr))

get(paste0(prefix, "_a_sig")) %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p) %>%
  dplyr::arrange(allele_dose_p)

a_sig_pos <- get(paste0(prefix, "_a_sig")) %>%
  dplyr::arrange(allele_dose_p) %>%
  dplyr::pull(position)

```

One for each SNP crossing significance threshold, arranged from most to least
significant. Only showing the top 6.

```{r out.width = '50%', fig.show = 'hold'}
# Inputs to plot_split should be
  # pheno_mfvc = pheno dataframe prepared above
  # "TopMed_ID" = name of ID column in pheno file
  # df_chr17 = dose frame dataframe prepared above
  # win_chr17_sig = significant LAAA results dataframe prepared above
  # .x = significant positions from win_chr17_sig prepared above
  # "maxFVC" = name of phenotype plotting - may need to update the plotting
    # function/units if changing this
  # colors = three plotting colors for manuscript
colors = c(color_lfvc, color_bfvc, color_maxfvc)

walk(a_sig_pos[1:6], ~ {
  print(plot_split(
    get(paste0("pheno_mfvc_", sex)),
    "TopMed_ID",
    get(paste0("df_chr", chr)),
    get(paste0(prefix, "_a_sig")), .x, "maxFVC", colors))
})
```

##### **Interaction Plot with Phenotype**

```{r}
get(paste0(prefix, "_int_sig")) %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p) %>%
  dplyr::arrange(allele_dose_p)

int_sig_pos <- get(paste0(prefix, "_int_sig")) %>%
  dplyr::arrange(allele_afr_dose_p) %>%
  dplyr::pull(position)
```


```{r out.width = '50%', fig.show = 'hold'}
walk(int_sig_pos[1:6], ~ {
  print(plot_split(
    get(paste0("pheno_mfvc_", sex)),
    "TopMed_ID",
    get(paste0("df_chr", chr)),
    get(paste0(prefix, "_int_sig")), .x, "maxFVC", colors))
})

```


## **MaxFEV1/FVC Chr10**

### **All Model P-Values**

```{r}
plot_sig(maxFEV1_FVC_chr10_female)
plot_sig(maxFEV1_FVC_chr10_male)
```




### **Miami Plots**

```{r out.width = '50%', fig.show = 'hold'}
plot_miami(maxFEV1_FVC_chr10_female, maxFEV1_FVC_chr10_male, "allele_dose_p")
plot_miami(maxFEV1_FVC_chr10_female, maxFEV1_FVC_chr10_male, "allele_afr_dose_p")
```

### **Allele & Interaction Results**

#### Female

##### **Allele Plot with Phenotype**

```{r}
# SET FOR WHOLE SECTION
pheno <- "maxFEV1_FVC"
chr <- "10"
sex <- "female"

# Get sig LAAA results & format
prefix <- paste0(pheno, "_chr", chr, "_", sex)
invisible(assign(
  paste0(prefix, "_a_sig"),
  get_sig(get(prefix), sig_thresh, "allele_dose_p")) %>%
  dplyr::mutate(chr = chr) %>%
  dplyr::arrange(allele_dose_p))
invisible(assign(
  paste0(prefix, "_int_sig"),
  get_sig(get(prefix), sig_thresh, "allele_afr_dose_p")) %>%
  dplyr::mutate(chr = chr) %>%
  dplyr::arrange(allele_afr_dose_p))

get(paste0(prefix, "_a_sig")) %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)

```

One for each SNP crossing significance threshold, arranged from most to least
significant. Only showing the top 6.

```{r out.width = '50%', fig.show = 'hold'}
# Inputs to plot_split should be
  # pheno_mfvc = pheno dataframe prepared above
  # "TopMed_ID" = name of ID column in pheno file
  # df_chr17 = dose frame dataframe prepared above
  # win_chr17_sig = significant LAAA results dataframe prepared above
  # .x = significant positions from win_chr17_sig prepared above
  # "maxFVC" = name of phenotype plotting - may need to update the plotting
    # function/units if changing this
  # colors = three plotting colors for manuscript
colors = c(color_lfvc, color_bfvc, color_maxfvc)


walk(get(paste0(prefix, "_a_sig"))$position, ~ {
  print(plot_split(
    get(paste0("pheno_mfvc_", sex)),
    "TopMed_ID",
    get(paste0("df_chr", chr)),
    get(paste0(prefix, "_a_sig")), .x, "maxFVC", colors))
})

```

##### **Interaction Plot with Phenotype**

```{r}
get(paste0(prefix, "_int_sig")) %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)
```


```{r out.width = '50%', fig.show = 'hold'}
walk(get(paste0(prefix, "_int_sig"))$position, ~ {
  print(plot_split(
    get(paste0("pheno_mfvc_", sex)),
    "TopMed_ID",
    get(paste0("df_chr", chr)),
    get(paste0(prefix, "_int_sig")), .x, "maxFVC", colors))
})

```


#### Male

##### **Allele Plot with Phenotype**

```{r}
# SET FOR WHOLE SECTION
pheno <- "maxFEV1_FVC"
chr <- "10"
sex <- "male"

# Get sig LAAA results & format
prefix <- paste0(pheno, "_chr", chr, "_", sex)
invisible(assign(
  paste0(prefix, "_a_sig"),
  get_sig(get(prefix), sig_thresh, "allele_dose_p")) %>%
  dplyr::mutate(chr = chr) %>%
  dplyr::arrange(allele_dose_p))
invisible(assign(
  paste0(prefix, "_int_sig"),
  get_sig(get(prefix), sig_thresh, "allele_afr_dose_p")) %>%
  dplyr::mutate(chr = chr) %>%
  dplyr::arrange(allele_afr_dose_p))

get(paste0(prefix, "_a_sig")) %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)

```

One for each SNP crossing significance threshold, arranged from most to least
significant. Only showing the top 6.

```{r out.width = '50%', fig.show = 'hold'}
# Inputs to plot_split should be
  # pheno_mfvc = pheno dataframe prepared above
  # "TopMed_ID" = name of ID column in pheno file
  # df_chr17 = dose frame dataframe prepared above
  # win_chr17_sig = significant LAAA results dataframe prepared above
  # .x = significant positions from win_chr17_sig prepared above
  # "maxFVC" = name of phenotype plotting - may need to update the plotting
    # function/units if changing this
  # colors = three plotting colors for manuscript
colors = c(color_lfvc, color_bfvc, color_maxfvc)

walk(get(paste0(prefix, "_a_sig"))$position, ~ {
  print(plot_split(
    get(paste0("pheno_mfvc_", sex)),
    "TopMed_ID",
    get(paste0("df_chr", chr)),
    get(paste0(prefix, "_a_sig")), .x, "maxFVC", colors))
})

```

##### **Interaction Plot with Phenotype**

```{r}
get(paste0(prefix, "_int_sig")) %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)
```


```{r out.width = '50%', fig.show = 'hold'}
walk(get(paste0(prefix, "_int_sig"))$position, ~ {
  print(plot_split(
    get(paste0("pheno_mfvc_", sex)),
    "TopMed_ID",
    get(paste0("df_chr", chr)),
    get(paste0(prefix, "_int_sig")), .x, "maxFVC", colors))
})

```


# **Write Out**

Want table of significant results for each chromosome/phenotype.

```{r}

```


# **Session Info**

```{r session_info}
sessionInfo()
```
