---
title: "3.2_laaa_replication"
author: "SDS"
date: "Last compiled on `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: kable
    code_folding: hide
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = paste0(Sys.getenv("RKJCOLLAB"), "/Collabs/ortega/reports")) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(
  root.dir = paste0(Sys.getenv("RKJCOLLAB"), "/Collabs/ortega"))

# Common packages
library(tidyverse)
library(knitr)
library(biomaRt)
library(DT)
library(conflicted)
library(scales)
library(patchwork)
library(table1)

# Set display options
options(digits = 4)  # only 2 digits after decimal

```

# **Overview**

Analysis of replication of SARP/CSGA LAAA results in BARD.

**TO NOTE: current report simplified to only look at chr8.**

# **Combined Analysis**
## **Load Data**

```{r load_data, message = F}
# TODO: Abhishek update the paths in this chunk to their locations on your computer
# SARP/CSGA LAAA significant results
data_sarp <- read_csv(
  "data/laaa_wgs/output/sig_all_results_all_chr.csv")
data_bard <- read_delim(
  "data/bard_laaa_results/adult_only/output/region_hg38_post_fvc_chr8/post_fvc_assoc_chr8.txt")

# Add max_abs_beta column to BARD data
data_bard$max_abs_beta <- apply(
    data_bard[, c("allele_dose_beta", "afr_dose_beta", "allele_afr_dose_beta")], 1, function(row) {
      cols <- c("allele_dose_beta", "afr_dose_beta", "allele_afr_dose_beta")
      cols[which.max(abs(row))]
    })

# Function to read in dose frames
read_dose_frames <- function(dir_name) {
    df_list <- list.files(dir_name, full.names = T)
    
    df_all <- Reduce(rbind, lapply(df_list, function(file) {
      file_base <- basename(file)
      df_name <- gsub(
        ".+[[:digit:]]+_", "",
        gsub("\\.txt", "", file_base))
      df <- read_delim(file) %>%
        dplyr::mutate(type = df_name) %>%
        relocate(type)
      return(df)
  }))
  return(df_all)
}

# Add paths to each directory that contains dose frames
df_chr8_sarp <-read_dose_frames(
  "data/laaa_wgs/dose_frames/region_hg38_maxFVC_chr8")
df_chr8_bard <- read_dose_frames(
  "data/bard_laaa_results/adult_only/dose_frames/region_hg38_post_fvc_chr8")

# Add paths to same phenotype files used while running LAAA
pheno_mfvc_sarp <- read_delim(
  "data/pheno/SARP123_CSGA_348_laaa_maxFVC_pheno.txt")
pheno_mfvc_bard <- read_delim(
  "data/bard_laaa_results/adult_only/pheno/BARD_age_gt_12_laaa_post_fvc_pheno.txt")

# Update path to script that sets main paper colors
source("/Users/slacksa/repos/local_ancestry_analysis/laaa_fine_mapping/set_main_paper_colors.R")

```

## **Functions**

```{r}
# Function to pivot longer and then plot
# data_df <- df_chr8_sarp
# data_af <- data_sarp
# positions <- data_sarp %>%
#   dplyr::filter(chr == "8") %>%
#   dplyr::pull(position)
# positions <- positions[1:10]

plot_dose_frames <- function(data_df, data_af, positions) {
  # Get list for factoring positions
  positions_fact <- as.character(sort(positions))
  
  # Afr dose
  data_df_mod <- data_df %>%
    dplyr::filter(position %in% positions) %>%
    pivot_longer(
      cols = -c("position", "ref", "alt", "type"),
      names_to = "ID",
      values_to = "value") %>%
  dplyr::mutate(
    value = factor(value, levels = c(0, 1, 2)),
    position = factor(position, levels = positions_fact),
    type = factor(type, levels = c("afr_dose", "allele_dose", "allele_afr_dose")))
  data_af_mod <- data_af %>%
    dplyr::filter(position %in% positions) %>%
    dplyr::mutate(
      position = factor(position, levels = positions_fact),
      int_na = as.factor(ifelse(is.na(allele_afr_dose_p), "yes", "no")))
  
  p1 <- ggplot(data_df_mod, aes(x = position, fill = value, group = interaction(type, value))) +
    geom_bar(position = "fill", width = 0.8) +
    theme_minimal() +
    facet_wrap(~type, ncol = 1) +
    theme(axis.text.x = element_blank(),
          axis.title.x = element_blank())
  p2 <- ggplot(data_af_mod, aes(x = position, y = alt_frq)) +
    geom_col() +
    theme_minimal() +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank())
  p3 <- ggplot(data_af_mod, aes(x = position, y = int_na, fill = int_na)) +
    geom_tile() +
    scale_fill_manual(values = c("no" = "grey", "yes" = "red")) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
          axis.ticks.x = element_blank())
  # Combined plot
  p1 / p2 / p3 + plot_layout(heights = c(4, 1, 1))
  
}

plot_dose_frames_spag <- function(data, positions) {
  # Afr dose
  data_mod <- data %>%
    dplyr::filter(position %in% positions) %>%
    pivot_longer(
      cols = -c("position", "ref", "alt", "type"),
      names_to = "ID",
      values_to = "value") %>%
  dplyr::mutate(
    value = as.numeric(value),  # keep 0, 1, 2 as numeric,
    position = as.factor(position),
    type = factor(type, levels = c("afr_dose", "allele_dose", "allele_afr_dose")))
  
    ggplot(data_mod, aes(x = position, y = value, group = ID)) +
      geom_line(alpha = 0.3) +              # line connecting each individual's values
      geom_point(alpha = 0.6, size = 1.5) + # dots at each position
      facet_wrap(~type, ncol = 1) +         # separate panel per dose type
      scale_y_continuous(breaks = c(0, 1, 2)) +
      labs(
        x = "Position",
        y = "Dose Value",
        title = "Per-Person Dose Trajectories Across Positions",
        subtitle = "Lines connect individual values per type"
      ) +
      theme_minimal() +
      theme(
        strip.text = element_text(face = "bold"),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
}

# Plot split phenotype
# For testing
# pheno_data <- pheno_mfvc_sarp
# pheno_data_id_col <- "TopMed_ID"
# dose_frame <- df_chr8_sarp
# sig_data <- data_sarp
# position <- data_sarp$position[1]
# pheno <- "maxFVC"
# colors = c(color_lfvc, color_bfvc, color_maxfvc)
# study <- "SARP"
fun_split <- function(x, data, pheno) {
  return(data.frame(
    y = min(data[[pheno]]) - (0.1*(max(data[[pheno]]) - min(data[[pheno]]))),
    label = length(x)))  # x comes from group = Asthma
}
plot_split <- function(
    pheno_data, pheno_data_id_col, dose_frame, sig_data, position, pheno, colors, study) {
 
  # Merge pheno and dose frame
  df_afr <- dose_frame %>% dplyr::filter(type == "afr_dose") %>%
    dplyr::select(-type, -ref, -alt) %>%
    column_to_rownames(var = "position")
  df_afr_t <- as.data.frame(t(df_afr))
  colnames(df_afr_t) <- paste0(colnames(df_afr_t), "_afr")
  df_afr_t <- rownames_to_column(df_afr_t, var = "ID")
  
  df_allele <- dose_frame %>% dplyr::filter(type == "allele_dose") %>%
    dplyr::select(-type, -ref, -alt) %>%
    column_to_rownames(var = "position")
  df_allele_t <- as.data.frame(t(df_allele))
  colnames(df_allele_t) <- paste0(colnames(df_allele_t), "_allele")
  df_allele_t <- rownames_to_column(df_allele_t, var = "ID")
  
  # Convert ID columns to chr before joining
  pheno_data[[pheno_data_id_col]] = as.character(pheno_data[[pheno_data_id_col]])
  df_afr_t$ID <- as.character(df_afr_t$ID)
  df_allele_t$ID <- as.character(df_allele_t$ID)
  
  pheno_df_temp <- inner_join(
    pheno_data,
    df_afr_t,
    by = setNames("ID", pheno_data_id_col))  # setNames flips order
  pheno_df <- inner_join(
    pheno_df_temp,
    df_allele_t,
    by = setNames("ID", pheno_data_id_col))  # setNames flips order
  
  # Get additional details for plot from sig_data
  driving_term <- sig_data$max_abs_beta[sig_data$position == position]
  
  # Make format of allele dose that have alt & ref alleles
  alt <- sig_data$alt[sig_data$position == position]
  ref <- sig_data$ref[sig_data$position == position]
  allele_labels <- c(
    `0` = paste0(ref, ref),
    `1` = paste0(ref, alt),
    `2` = paste0(alt, alt))
  
  plot <- ggplot(pheno_df, aes(
    x = factor(!!rlang::sym(paste0(position, "_allele"))),
    y = !!rlang::sym(pheno))) +
  geom_boxplot(aes(fill = "Combined"),
               varwidth = TRUE, alpha = 0.7, outlier.shape = 21) +
  geom_boxplot(aes(fill = factor(!!rlang::sym(paste0(position, "_afr")))),
               varwidth = TRUE, alpha = 0.7, outlier.shape = 21) +
  labs(x = "Genotype",
       y = paste0(
         gsub("_", "/", gsub("m", "M", pheno)),
         ifelse(str_detect(pheno, "_"), " Ratio", " (mL)")),
       fill = "Copies Local African Ancestry",
       title = paste0(
         study, " - chr", sig_data$chr, ":", position, ":", ref, ":", alt, ", Driving Term: ", driving_term)) +
  scale_x_discrete(labels = allele_labels) +
  scale_y_continuous(labels = label_comma()) +
  scale_fill_manual(values = c("Combined" = "transparent",  # Transparent box plot
                             "0" = colors[1], "1" = colors[2], "2" = colors[3]),
                    labels = c("Combined" = "Combined", "0" = "0",
                               "1" = "1", "2" = "2")) +
  theme_bw() +
  theme(legend.position = "bottom") +
  stat_summary(fun.data = fun_split, fun.args = list(data = pheno_df, pheno = pheno),
               geom = "text", aes(group = !!rlang::sym(paste0(position, "_afr"))),
               position = position_dodge(0.5), hjust = 0.5)
  return(plot)
}

```

## **Compare**

For comparison - first getting significant LAAA results in SARP/CSGA, then
subsetting to only those SNPs that are also found in BARD results.

SARP/CSGA covariates: group, age, sex, ht, bmi, RFMIX_GW_AFR

BARD covariates: age_screening, gender, bmi_screening, height_cm, RFMIX_GW_AFR

*TODO: we may want to consider removing covariates not significantly associated*
*with max/post FVC.*

## **Table 1s**

These are only the individuals included in the maxFVC phenotype files for LAAA.

```{r}
pheno_mfvc_merge <- full_join(
  pheno_mfvc_sarp %>%
    dplyr::mutate(
      study = "SARP_CSGA"),
  pheno_mfvc_bard %>% dplyr::mutate(
    FID = as.character(FID),
    IID = as.character(IID),
    study = "BARD"),
  by = c("FID", "IID", "RFMIX_GW_AFR", "study", "maxFVC" = "post_fvc",
         "ht" = "height_cm", "sex" = "gender", "age" = "age_screening",
         "bmi" = "bmi_screening"))


table1(~ maxFVC + ht + age + as.factor(sex) + as.factor(group) + bmi +
         RFMIX_GW_AFR + weight_kg | study,
       data = pheno_mfvc_merge)
```


## **MaxFVC Chr8**


### **Dose Frames**

Trying to understand the many NA values for interaction (allele_afr_dose) in 
BARD. In Lancet Child Adolesc Health paper, Michelle comments that this often
occurs for lower frequency SNPs because they are more likely to be on a single
ancestral background, which leads to no variability in allele_afr_dose.

**SARP/CSGA**

```{r}
sig_pos_chr8 <- data_sarp %>%
  dplyr::filter(chr == "8") %>%
  dplyr::filter(position %in% data_bard$position) %>%
  dplyr::pull(position)

plot_dose_frames(df_chr8_sarp, data_sarp, sig_pos_chr8)
```

**BARD**

```{r}
plot_dose_frames(df_chr8_bard, data_bard, sig_pos_chr8)
```


Across the region, the African ancestry dose hardly changes at all in either
study. The two studies are incredibly consistent! Is sample size driving the 
increase in interaction NAs in BARD?

Check if individuals change values in this region:

*TODO: update this with better way to keep individuals separate?*

**SARP/CSGA**

```{r}
plot_dose_frames_spag(df_chr8_sarp, sig_pos_chr8)
```

**BARD**

```{r}
plot_dose_frames_spag(df_chr8_bard, sig_pos_chr8)
```


Alternate allele frequencies look nearly identical - quick check that they are
not:

```{r}
check_af <- inner_join(
  data_bard %>% dplyr::select(position, alt_frq),
  data_sarp %>% dplyr::select(position, alt_frq),
  by = c("position"),
  suffix = c("_b", "_s"))

datatable(check_af)
```


### **Phenotype Plots**

For list of SNPs with NAs, look at phenotype plots side by side in BARD &
SARP/CSGA.

**BARD is on the left, SARP/CSGA is on the right.**


```{r out.width = '50%', fig.show = 'hold'}
sig_pos_chr8_bard_na <- data_bard %>%
  dplyr::filter(position %in% sig_pos_chr8) %>%
  dplyr::filter(is.na(allele_afr_dose_p)) %>%
  dplyr::pull(position)

colors = c(color_lfvc, color_bfvc, color_maxfvc)

data_bard$chr <- 8
# plot_split(pheno_mfvc_bard, "IID", df_chr8_bard, data_bard, "53887100", "post_fvc", colors)
# plot_split(pheno_mfvc_sarp, "TopMed_ID", df_chr8_sarp, data_sarp, "53887100", "maxFVC", colors)

walk(sig_pos_chr8_bard_na, ~ {
  print(plot_split(pheno_mfvc_bard, "IID", df_chr8_bard, data_bard, .x, "post_fvc", colors, "BARD"))
  print(plot_split(pheno_mfvc_sarp, "TopMed_ID", df_chr8_sarp, data_sarp, .x, "maxFVC", colors, "SARP/CSGA"))
})
```



## **For Manuscript**

For first version of manuscript, including boxplots for the top replicated SNP,
instead of the top SNP in SARP/CSGA

```{r}
data_merge <- inner_join(
  data_bard,
  data_sarp,
  by = c("chr", "position"),
  suffix = c("_b", "_s")) %>%
  dplyr::select(
    position, ref_b, alt_b, alt_frq_b, ref_s, alt_s, alt_frq_s,
    allele_dose_p_b, allele_dose_p_s)

head(data_merge[order(data_merge$allele_dose_p_s), ])
```

```{r}
head(data_merge[order(data_merge$allele_dose_p_b), ])
```

First BARD SNP with p < 0.05 is at position 53917720. Make boxplots in SARP/CSGA
and in BARD for this SNP.

```{r}
plot_split(
  pheno_mfvc_bard, "IID", df_chr8_bard, data_bard, "53917720",
  "post_fvc", colors, "BARD")
plot_split(
  pheno_mfvc_sarp, "TopMed_ID", df_chr8_sarp, data_sarp, "53917720",
  "maxFVC", colors, "SARP/CSGA")
```



# **Stratified by Sex Analysis**

Plan is to look at the sex-stratified admixture mapping significant regions in
both sexes in SARP & BARD, regardless of which sex the admixture mapping peak
was identified in. Will do this with SARP as primary analysis, so for
significant LAAA results in SARP (p-allele or p-afr-allele < 0.001), will merge
in BARD results, then see if any BARD results also significant (< 0.05). Will
export 1. full SARP results with BARD merged in & 2. only rows significant in
both.

## **Load Data**

```{r load_data, message = F}


data_s <- read_delim(
  "data/laaa_wgs/strat_by_sex/output_male/region_hg38_maxFVC_chr8/maxFVC_assoc_chr8.txt")
data_b <- read_delim(
  "data/bard_laaa_results/adult_only/strat_by_sex/laaa_wgs_male/output/region_hg38_post_fvc_chr8/post_fvc_assoc_chr8.txt")


```

## **Tables LAAA Result Overlap**

Looking at SNPs with p allele dose or p allele_afr dose < 10^-3 in SARP/CSGA,
and pulling the results from those SNPs in BARD.

### MaxFVC chr8 Males


```{r}
female_s_list <- c(
  "chr5_maxFVC" = "data/laaa_wgs/strat_by_sex/output_female/region_hg38_maxFVC_chr5/maxFVC_assoc_chr5.txt",
  "chr8_maxFVC" = "data/laaa_wgs/strat_by_sex/output_female/region_hg38_maxFVC_chr8/maxFVC_assoc_chr8.txt",
  "chr1_maxFEVFVC" = "data/laaa_wgs/strat_by_sex/output_female/region_hg38_maxFEV1_FVC_chr1/maxFEV1_FVC_assoc_chr1.txt",
  "chr6_maxFEVFVC" = "data/laaa_wgs/strat_by_sex/output_female/region_hg38_maxFEV1_FVC_chr6/maxFEV1_FVC_assoc_chr6.txt",
  "chr10_maxFEVFVC" = "data/laaa_wgs/strat_by_sex/output_female/region_hg38_maxFEV1_FVC_chr10/maxFEV1_FVC_assoc_chr10.txt",
  "chr16_maxFEVFVC" = "data/laaa_wgs/strat_by_sex/output_female/region_hg38_maxFEV1_FVC_chr16/maxFEV1_FVC_assoc_chr16.txt",
  "chr1_maxFEV" = "data/laaa_wgs/strat_by_sex/output_female/region_hg38_maxFEV1_chr1/maxFEV1_assoc_chr1.txt",
  "chr2_bFEVFVC" = "data/laaa_wgs/strat_by_sex/output_female/region_hg38_bFEV_FVC_chr2/bFEV_FVC_assoc_chr2.txt",
  "chr3_bFEVFVC" = "data/laaa_wgs/strat_by_sex/output_female/region_hg38_bFEV_FVC_chr3/bFEV_FVC_assoc_chr3.txt")

female_b_list <- c(
  "chr5_maxFVC" = "data/bard_laaa_results/adult_only/strat_by_sex/laaa_wgs_female/output/region_hg38_post_fvc_chr5/post_fvc_assoc_chr5.txt",
  "chr8_maxFVC" = "data/bard_laaa_results/adult_only/strat_by_sex/laaa_wgs_female/output/region_hg38_post_fvc_chr8/post_fvc_assoc_chr8.txt",
  "chr1_maxFEVFVC" = "data/bard_laaa_results/adult_only/strat_by_sex/laaa_wgs_female/output/region_hg38_post_fev1fvc_ratio_chr1/post_fev1fvc_ratio_assoc_chr1.txt",
  "chr6_maxFEVFVC" = "data/bard_laaa_results/adult_only/strat_by_sex/laaa_wgs_female/output/region_hg38_post_fev1fvc_ratio_chr6/post_fev1fvc_ratio_assoc_chr6.txt",
  "chr10_maxFEVFVC" = "data/bard_laaa_results/adult_only/strat_by_sex/laaa_wgs_female/output/region_hg38_post_fev1fvc_ratio_chr10/post_fev1fvc_ratio_assoc_chr10.txt",
  "chr16_maxFEVFVC" = "data/bard_laaa_results/adult_only/strat_by_sex/laaa_wgs_female/output/region_hg38_post_fev1fvc_ratio_chr16/post_fev1fvc_ratio_assoc_chr16.txt",
  "chr1_maxFEV" = "data/bard_laaa_results/adult_only/strat_by_sex/laaa_wgs_female/output/region_hg38_post_fev1_chr1/post_fev1_assoc_chr1.txt",
  "chr2_bFEVFVC" = "data/bard_laaa_results/adult_only/strat_by_sex/laaa_wgs_female/output/region_hg38_bfev1fvc_ratio_chr2/bfev1fvc_ratio_assoc_chr2.txt",
  "chr3_bFEVFVC" = "data/bard_laaa_results/adult_only/strat_by_sex/laaa_wgs_female/output/region_hg38_bfev1fvc_ratio_chr3/bfev1fvc_ratio_assoc_chr3.txt")

male_s_list <- c(
  "chr5_maxFVC" = "data/laaa_wgs/strat_by_sex/output_male/region_hg38_maxFVC_chr5/maxFVC_assoc_chr5.txt",
  "chr8_maxFVC" = "data/laaa_wgs/strat_by_sex/output_male/region_hg38_maxFVC_chr8/maxFVC_assoc_chr8.txt",
  "chr1_maxFEVFVC" = "data/laaa_wgs/strat_by_sex/output_male/region_hg38_maxFEV1_FVC_chr1/maxFEV1_FVC_assoc_chr1.txt",
  "chr6_maxFEVFVC" = "data/laaa_wgs/strat_by_sex/output_male/region_hg38_maxFEV1_FVC_chr6/maxFEV1_FVC_assoc_chr6.txt",
  "chr10_maxFEVFVC" = "data/laaa_wgs/strat_by_sex/output_male/region_hg38_maxFEV1_FVC_chr10/maxFEV1_FVC_assoc_chr10.txt",
  "chr16_maxFEVFVC" = "data/laaa_wgs/strat_by_sex/output_male/region_hg38_maxFEV1_FVC_chr16/maxFEV1_FVC_assoc_chr16.txt",
  "chr1_maxFEV" = "data/laaa_wgs/strat_by_sex/output_male/region_hg38_maxFEV1_chr1/maxFEV1_assoc_chr1.txt",
  "chr2_bFEVFVC" = "data/laaa_wgs/strat_by_sex/output_male/region_hg38_bFEV_FVC_chr2/bFEV_FVC_assoc_chr2.txt",
  "chr3_bFEVFVC" = "data/laaa_wgs/strat_by_sex/output_male/region_hg38_bFEV_FVC_chr3/bFEV_FVC_assoc_chr3.txt")

male_b_list <- c(
  "chr5_maxFVC" = "data/bard_laaa_results/adult_only/strat_by_sex/laaa_wgs_male/output/region_hg38_post_fvc_chr5/post_fvc_assoc_chr5.txt",
  "chr8_maxFVC" = "data/bard_laaa_results/adult_only/strat_by_sex/laaa_wgs_male/output/region_hg38_post_fvc_chr8/post_fvc_assoc_chr8.txt",
  "chr1_maxFEVFVC" = "data/bard_laaa_results/adult_only/strat_by_sex/laaa_wgs_male/output/region_hg38_post_fev1fvc_ratio_chr1/post_fev1fvc_ratio_assoc_chr1.txt",
  "chr6_maxFEVFVC" = "data/bard_laaa_results/adult_only/strat_by_sex/laaa_wgs_male/output/region_hg38_post_fev1fvc_ratio_chr6/post_fev1fvc_ratio_assoc_chr6.txt",
  "chr10_maxFEVFVC" = "data/bard_laaa_results/adult_only/strat_by_sex/laaa_wgs_male/output/region_hg38_post_fev1fvc_ratio_chr10/post_fev1fvc_ratio_assoc_chr10.txt",
  "chr16_maxFEVFVC" = "data/bard_laaa_results/adult_only/strat_by_sex/laaa_wgs_male/output/region_hg38_post_fev1fvc_ratio_chr16/post_fev1fvc_ratio_assoc_chr16.txt",
  "chr1_maxFEV" = "data/bard_laaa_results/adult_only/strat_by_sex/laaa_wgs_male/output/region_hg38_post_fev1_chr1/post_fev1_assoc_chr1.txt",
  "chr2_bFEVFVC" = "data/bard_laaa_results/adult_only/strat_by_sex/laaa_wgs_male/output/region_hg38_bfev1fvc_ratio_chr2/bfev1fvc_ratio_assoc_chr2.txt",
  "chr3_bFEVFVC" = "data/bard_laaa_results/adult_only/strat_by_sex/laaa_wgs_male/output/region_hg38_bfev1fvc_ratio_chr3/bfev1fvc_ratio_assoc_chr3.txt")

for (region in names(female_s_list)) {  # can use any list, all same names
  print(region)
  for (sex in c("female", "male")) {
    print(sex)
    s_list <- get(paste0(sex, "_s_list"))
    b_list <- get(paste0(sex, "_b_list"))

    data_s <- read_delim(s_list[region], show_col_types = F)
    data_b <- read_delim(b_list[region], show_col_types = F)
    
    data_s_sig <- data_s %>%
      dplyr::filter(allele_dose_p < 10^-3 | allele_afr_dose_p < 10^-3)
  
    merge <- left_join(
      data_s_sig,
      data_b,
      by = c("position", "ref", "alt"),
      suffix = c("_s", "_b"))
    
    merge_exp <- merge %>%
      dplyr::select(
        position, ref, alt, alt_frq_s, alt_frq_b,
        all_of(starts_with("allele")), all_of(starts_with("afr"))) %>%
      dplyr::select(
        position, ref, alt, alt_frq_s, alt_frq_b,
        allele_dose_beta_s, allele_dose_p_s, allele_dose_beta_b, allele_dose_p_b,
        allele_afr_dose_beta_s, allele_afr_dose_p_s, allele_afr_dose_beta_b, allele_afr_dose_p_b,
        afr_dose_beta_s, afr_dose_p_s, afr_dose_beta_b, afr_dose_p_b) %>%
      dplyr::mutate(
        sig_allele_dose_p_s = ifelse(allele_dose_p_s < 10^-3, 1, 0),
        sig_allele_afr_dose_p_s = ifelse(allele_afr_dose_p_s < 10^-3, 1, 0))
    print("SARP-CSGA N sig:")
    print(nrow(merge_exp))
    
    merge_exp_sig <- merge_exp %>%
      dplyr::filter(
        (sig_allele_dose_p_s ==1 & allele_dose_p_b < 0.05) |
          (sig_allele_afr_dose_p_s == 1 & allele_afr_dose_p_b < 0.05))
    print("SARP-CSGA & BARD N sig:")
    print(nrow(merge_exp_sig))
    
    # write_tsv(
    #   merge_exp,
    #   paste0("data/bard_laaa_results/adult_only/strat_by_sex/comp_to_sarp_csga/",
    #          region, "_", sex, "_all_sarp_csga_sig.txt"))
    # write_tsv(
    #   merge_exp_sig,
    #   paste0("data/bard_laaa_results/adult_only/strat_by_sex/comp_to_sarp_csga/",
    #          region, "_", sex, "_sarp_csga_and_bard_sig.txt"))
  }
}
```


# **Session Info**

```{r session_info}
sessionInfo()
```
