---
title: "3_laaa_results"
author: "SDS"
date: "Last compiled on `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: kable
    code_folding: hide
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = paste0(Sys.getenv("RKJCOLLAB"), "/Collabs/ortega/reports")) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(
  root.dir = paste0(Sys.getenv("RKJCOLLAB"), "/Collabs/ortega"))

# Common packages
library(tidyverse)
library(knitr)
library(biomaRt)
library(DT)
library(conflicted)
library(scales)

#  For locuszoomr
library(locuszoomr)
library(ensembldb)
library(AnnotationHub)
library(GGally)
library(RColorBrewer)
library(ggpubr)
library(IRanges)
library(snpStats)

# Set display options
options(digits = 4)  # only 2 digits after decimal

```

# **Overview**

Makes combined locus zoom and phenotype plots from LAAA/fine mapping results.


**LAAA model formula:**

pheno ~ group + covariates + RFMIX_GW_AFR + allele_dose + afr_dose + allele_afr_dose

For locus zoom plots, LD is calculated within our study population using
R package snpStats.

Significance threshold can be changed, but current script follows Dr. Ortega's
and Michelle's previous paper: P < 10^-3 (0.001). Significant SNPs are those
whose allele dose p-value and/or interaction dose p-value meets this threshold.

# **Load Data**

```{r load_data, message = F}

# TODO: update all paths in this section

# Add paths to each set of LAAA analyses run, with results in hg38
win_chr5 <- read_delim(
  "data/laaa_wgs/output/region_hg38_maxFVC_chr5/maxFVC_assoc_chr5.txt")
win_chr8 <- read_delim(
  "data/laaa_wgs/output/region_hg38_maxFVC_chr8/maxFVC_assoc_chr8.txt")
win_chr10 <- read_delim(
  "data/laaa_wgs/output/region_hg38_maxFEV1_FVC_chr10/maxFEV1_FVC_assoc_chr10.txt")
win_chr17 <- read_delim(
  "data/laaa_wgs/output/region_hg38_maxFVC_chr17/maxFVC_assoc_chr17.txt")

# Function to read in dose frames
read_dose_frames <- function(dir_name) {
    df_list <- list.files(dir_name, full.names = T)
    
    df_all <- Reduce(rbind, lapply(df_list, function(file) {
      file_base <- basename(file)
      df_name <- gsub(
        ".+[[:digit:]]+_", "",
        gsub("\\.txt", "", file_base))
      df <- read_delim(file) %>%
        dplyr::mutate(type = df_name) %>%
        relocate(type)
      return(df)
  }))
  return(df_all)
}

# Add paths to each directory that contains dose frames
df_chr5 <-
  read_dose_frames("data/laaa_wgs/dose_frames/region_hg38_maxFVC_chr5")
df_chr8 <-
  read_dose_frames("data/laaa_wgs/dose_frames/region_hg38_maxFVC_chr8")
df_chr10 <-
  read_dose_frames("data/laaa_wgs/dose_frames/region_hg38_maxFEV1_FVC_chr10")
df_chr17 <-
  read_dose_frames("data/laaa_wgs/dose_frames/region_hg38_maxFVC_chr17")

# Add paths to same phenotype files used while running LAAA
pheno_mff <- read_delim(
  "data/pheno/SARP123_CSGA_348_laaa_maxFEV1_FVC_pheno.txt")
pheno_mfvc <- read_delim(
  "data/pheno/SARP123_CSGA_348_laaa_maxFVC_pheno.txt")

# Update path to script that sets main paper colors
source("/Users/slacksa/repos/local_ancestry_analysis/laaa_fine_mapping/set_main_paper_colors.R")

# Set significance thresholds - both normal and nominal
sig_thresh <- 0.001
nom_sig_thresh <- 0.01

```

**Regions Tested:**

+ MaxFVC chr5:`r format(min(win_chr5$position), scientific = F)` - `r format(max(win_chr5$position), scientific = F)`
+ MaxFVC chr8:`r format(min(win_chr8$position), scientific = F)` - `r format(max(win_chr8$position), scientific = F)`
+ MaxFVC chr17:`r format(min(win_chr17$position), scientific = F)` - `r format(max(win_chr17$position), scientific = F)`
+ MaxFEV1/FVC chr10:`r format(min(win_chr10$position), scientific = F)` - `r format(max(win_chr10$position), scientific = F)`

# **Functions**

```{r}
# Get significant results
# Can run to get significant for given p-value
get_sig <- function(df, thresh, pval) {
  # Get significant results
  df_sig <- df %>% dplyr::filter(!!sym(pval) < thresh)
  
  # Write abs(max) column name in new column
  df_sig$max_abs_beta <- apply(
    df_sig[, c("allele_dose_beta", "afr_dose_beta", "allele_afr_dose_beta")],
    1, function(row) {
      cols <- c("allele_dose_beta", "afr_dose_beta", "allele_afr_dose_beta")
      cols[which.max(abs(row))]
  })
  return(df_sig)
}

# Make plotting function - all p-values in one plot
plot_sig <- function(win_df) {

  win_df_lg <- win_df %>%
    dplyr::select(position, allele_dose_p, afr_dose_p, allele_afr_dose_p, anova_p) %>%
    pivot_longer(cols = c(allele_dose_p, afr_dose_p, allele_afr_dose_p, anova_p),
                 names_to = "type",
                 values_to = "pval")
  
  win_df_sig <- subset(win_df_lg, pval < sig_thresh & type == "anova_p")
  win_df_nom_sig <- subset(win_df_lg, pval < nom_sig_thresh & type == "anova_p")
  
  ggplot(win_df_lg, aes(position, -log10(pval))) +
    geom_point() +
    geom_point(data = subset(win_df_lg, position %in% win_df_sig$position),
               aes(color = "orange")) +
    labs(x = "Position", y = "-log10(P-Value)", color = paste0("Anova P < ", sig_thresh)) +
    facet_wrap(~ type) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

# Plot split phenotype
# For testing
# pheno_data <- pheno_mfvc
# pheno_data_id_col <- "TopMed_ID"
# dose_frame <- df_chr8
# sig_data <- win_chr8_a_sig
# position <- win_chr8_a_sig_peaks$position[1]
# pheno <- "maxFVC"
# colors = c(color_lfvc, color_bfvc, color_maxfvc)
fun_split <- function(x, data, pheno) {
  return(data.frame(
    y = min(data[[pheno]]) - (0.1*(max(data[[pheno]]) - min(data[[pheno]]))),
    label = length(x)))  # x comes from group = Asthma
}
plot_split <- function(
    pheno_data, pheno_data_id_col, dose_frame, sig_data, position, pheno, colors) {
 
  # Merge pheno and dose frame
  df_afr <- dose_frame %>% dplyr::filter(type == "afr_dose") %>%
    dplyr::select(-type, -ref, -alt) %>%
    column_to_rownames(var = "position")
  df_afr_t <- as.data.frame(t(df_afr))
  colnames(df_afr_t) <- paste0(colnames(df_afr_t), "_afr")
  df_afr_t <- rownames_to_column(df_afr_t, var = "ID")
  
  df_allele <- dose_frame %>% dplyr::filter(type == "allele_dose") %>%
    dplyr::select(-type, -ref, -alt) %>%
    column_to_rownames(var = "position")
  df_allele_t <- as.data.frame(t(df_allele))
  colnames(df_allele_t) <- paste0(colnames(df_allele_t), "_allele")
  df_allele_t <- rownames_to_column(df_allele_t, var = "ID")
  
  # Convert ID columns to chr before joining
  pheno_data[[pheno_data_id_col]] = as.character(pheno_data[[pheno_data_id_col]])
  df_afr_t$ID <- as.character(df_afr_t$ID)
  df_allele_t$ID <- as.character(df_allele_t$ID)
  
  pheno_df_temp <- inner_join(
    pheno_data,
    df_afr_t,
    by = setNames("ID", pheno_data_id_col))  # setNames flips order
  pheno_df <- inner_join(
    pheno_df_temp,
    df_allele_t,
    by = setNames("ID", pheno_data_id_col))  # setNames flips order
  
  # Get additional details for plot from sig_data
  driving_term <- sig_data$max_abs_beta[sig_data$position == position]
  
  # Make format of allele dose that have alt & ref alleles
  alt <- sig_data$alt[sig_data$position == position]
  ref <- sig_data$ref[sig_data$position == position]
  allele_labels <- c(
    `0` = paste0(ref, ref),
    `1` = paste0(ref, alt),
    `2` = paste0(alt, alt))
  
  plot <- ggplot(pheno_df, aes(
    x = factor(!!rlang::sym(paste0(position, "_allele"))),
    y = !!rlang::sym(pheno))) +
  geom_boxplot(aes(fill = "Combined"),
               varwidth = TRUE, alpha = 0.7, outlier.shape = 21) +
  geom_boxplot(aes(fill = factor(!!rlang::sym(paste0(position, "_afr")))),
               varwidth = TRUE, alpha = 0.7, outlier.shape = 21) +
  labs(x = "Genotype",
       y = paste0(
         gsub("_", "/", gsub("m", "M", pheno)),
         ifelse(str_detect(pheno, "_"), " Ratio", " (mL)")),
       fill = "Copies Local African Ancestry",
       title = paste0(
         "chr", sig_data$chr, ":", position, ":", ref, ":", alt, ", Driving Term: ", driving_term)) +
  scale_x_discrete(labels = allele_labels) +
  scale_y_continuous(labels = label_comma()) +
  scale_fill_manual(values = c("Combined" = "transparent",  # Transparent box plot
                             "0" = colors[1], "1" = colors[2], "2" = colors[3]),
                    labels = c("Combined" = "Combined", "0" = "0",
                               "1" = "1", "2" = "2")) +
  theme_bw() +
  theme(legend.position = "bottom") +
  stat_summary(fun.data = fun_split, fun.args = list(data = pheno_df, pheno = pheno),
               geom = "text", aes(group = !!rlang::sym(paste0(position, "_afr"))),
               position = position_dodge(0.5), hjust = 0.5)
  return(plot)
}

# Get LD
# peak_snp <- win_chr5_a_sig_peaks$position[1]
# win_sig <- win_chr5_a_sig
# dose_frame <- df_chr5
# pval <- "allele_dose_p"
calc_peak_ld <- function(peak_snp, win_sig, dose_frame, pval) {
  # Get peak snp
  peak <- paste0(
    win_sig$position[which.min(win_sig[[pval]])], "_allele")
  
  # Modify dose frame, use to create SnpMatrix
  df_allele_t <- as.data.frame(t(dose_frame %>%
    dplyr::filter(type == "allele_dose") %>%
    dplyr::select(-type, -ref, -alt) %>%
    column_to_rownames(var = "position")))
  snp_m <- new("SnpMatrix", as.matrix(df_allele_t))
  
  # Get pairwise LD with peak SNP
  ld <- ld(snp_m[, which(colnames(snp_m) == peak_snp), drop = F],
           snp_m, stats = c("R.squared"))
  
  # LD R2 column name is just position of peak SNP calculated with respect to
  ld_t <- as.data.frame(t(ld)) %>%
    rownames_to_column("pos") %>%
    dplyr::mutate(pos = as.integer(gsub("_allele", "", pos)))
  
  return(ld_t)
}

# Plot locus zoom
# win_ld <- win_chr8_int_ld
# chr <- "8"
# peak_snp <- chr8_int_peak
# pval <- "allele_afr_dose_p"
# win_ld <- win_chr5_a_ld
# chr <- "5"
# peak_snp <- chr5_a_peak
# pval <- "allele_dose_p"
plot_locuszoom <- function(win_ld, chr, peak_snp, pval) {
  # Remove NA pvalues
  # TODO: revisit and confirm this as expected!
  win_ld <- win_ld[!is.na(win_ld[[pval]]), ]
  # Plot
  loc <- locus(data = win_ld, chrom = "chr", p = pval, pos = "position",
             ens_db = ensDb_v111, seqname = chr, labs = "position",
             xrange = c(min(win_ld$position), max(win_ld$position)),
             LD = as.character(peak_snp))
  loc_plot <- locus_ggplot(loc, filter_gene_biotype = "protein_coding",
             pcutoff = sig_thresh, legend_pos = "top")
  return(loc_plot)
}

```


# **Results**

For each window, first check for overall model significance. For each variant
that is significant, then identify the effect with the largest absolute value
of the test statistics to determine which part of the LAAA model is driving
that significance.

## **MaxFVC Chr5**

### **All Model P-Values**

```{r}
plot_sig(win_chr5)
```

Based on this, think makes sense to only look at allele dose results and
interaction results below - both anova of full model compared to null model and
African ancestry dose results are very inflated.

### **Allele & Interaction Results**

Meeting nominal significance threshold of: `r sig_thresh`.

Allele:

```{r}
# Get sig LAAA results & format
win_chr5_a_sig <- get_sig(win_chr5, sig_thresh, "allele_dose_p")
win_chr5_int_sig <- get_sig(win_chr5, sig_thresh, "allele_afr_dose_p")


win_chr5_a_sig %>%
  dplyr::arrange(allele_dose_p) %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)

```



Interaction:

```{r}
win_chr5_int_sig %>%
  dplyr::arrange(allele_afr_dose_p) %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)
```


## **MaxFVC Chr8**

### **All Model P-Values**

```{r}
plot_sig(win_chr8)
```

Based on this, think makes sense to only look at allele dose results and
interaction results below - both anova of full model compared to null model and
African ancestry dose results are very inflated.

### **Allele & Interaction Results**

Meeting significance threshold of: `r sig_thresh`.

Allele:

```{r}
# Get sig LAAA results & format
win_chr8_a_sig <- get_sig(win_chr8, sig_thresh, "allele_dose_p")
win_chr8_int_sig <- get_sig(win_chr8, sig_thresh, "allele_afr_dose_p")

win_chr8_a_sig %>%
  dplyr::arrange(allele_dose_p) %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)

```


Interaction:

```{r}
win_chr8_int_sig %>%
  dplyr::arrange(allele_afr_dose_p) %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)
```


Combined:

```{r}
win_chr8_sig_merge <- full_join(
  win_chr8_a_sig %>% dplyr::mutate(allele_dose_p_sig = 1),
  win_chr8_int_sig %>% dplyr::mutate(allele_afr_dose_p_sig = 1),
  by = c(colnames(win_chr8_a_sig))) %>%
  dplyr::mutate(
    allele_dose_p_sig = ifelse(is.na(allele_dose_p_sig), 0, allele_dose_p_sig),
    allele_afr_dose_p_sig = ifelse(is.na(allele_afr_dose_p_sig), 0, allele_afr_dose_p_sig))

win_chr8_sig_merge %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)
```

Combined - summarized to keep only the first of those in complete LD:

For now, doing this by filtering to keep only one of multiple rows with
identical alt_frq, anova_p, max_abs_beta, and allele or interaction significance.

```{r}
win_chr8_sig_merge_dedup <- win_chr8_sig_merge %>%
  distinct(alt_frq, anova_p, max_abs_beta, allele_dose_p_sig,
           allele_afr_dose_p_sig, .keep_all = TRUE)

win_chr8_sig_merge_dedup %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)

```


### **Allele Locus Zoom Plot**

*TODO: will revisit adding these later - want to only calculated LD once*

For locus zoom plot, want one version for each of the "peaks". To do this,
narrow down the above to keep only top SNP from each set of physically close
SNPs.

```{r}
# Get ranges of points within 30,000bp of each other
win_chr8_a_sig_ranges <-
  IRanges(start = win_chr8_a_sig$position, end = win_chr8_a_sig$position)
win_chr8_a_sig_merged_ranges <-
  IRanges::reduce(win_chr8_a_sig_ranges, min.gapwidth = 30000)

# Add range group to original df
win_chr8_a_sig$group <-
  subjectHits(findOverlaps(win_chr8_a_sig_ranges, win_chr8_a_sig_merged_ranges))

# Group by the merged ranges and keep the row with the lowest p-value in each group
win_chr8_a_sig_peaks <- win_chr8_a_sig %>%
  dplyr::group_by(group) %>%
  dplyr::slice(which.min(allele_dose_p)) %>%
  ungroup() %>%
  arrange(allele_dose_p)

head(win_chr8_a_sig_peaks)
```

Only plot the one most significant peak SNP:

```{r}
# chr8_a_ld <- Reduce(full_join, lapply(win_chr8_a_sig_peaks$position, function(pos) {
#   return(calc_peak_ld(pos, win_chr8_a_sig, df_chr8, "allele_dose_p"))
# }))
# 
# win_chr8_a_ld <- as.data.frame(inner_join(
#   win_chr8 %>% dplyr::mutate(chr = 8),
#   chr8_a_ld,
#   by = c("position" = "pos")))  # should keep all entries
# 
# 
# chr8_a_peak <- win_chr8_a_sig_peaks$position[which.min(win_chr8_a_sig_peaks$allele_dose_p)]
# 
# # Plot locus zoom
# plot_chr8_a_lz <- plot_locuszoom(win_chr8_a_ld, "8", chr8_a_peak, "allele_dose_p")
# plot_chr8_a_lz

```


### **Interaction Locus Zoom Plot**

For locus zoom plot, want one version for each of the "peaks". To do this,
narrow down the above to keep only top SNP from each set of physically close
SNPs.

```{r}
# Get ranges of points within 30,000bp of each other
win_chr8_int_sig_ranges <-
  IRanges(start = win_chr8_int_sig$position, end = win_chr8_int_sig$position)
win_chr8_int_sig_merged_ranges <-
  IRanges::reduce(win_chr8_int_sig_ranges, min.gapwidth = 30000)

# Add range group to original df
win_chr8_int_sig$group <-
  subjectHits(findOverlaps(win_chr8_int_sig_ranges, win_chr8_int_sig_merged_ranges))

# Group by the merged ranges and keep the row with the lowest p-value in each group
win_chr8_int_sig_peaks <- win_chr8_int_sig %>%
  dplyr::group_by(group) %>%
  dplyr::slice(which.min(allele_afr_dose_p)) %>%
  ungroup() %>%
  arrange(allele_afr_dose_p)

head(win_chr8_int_sig_peaks)
```

Only plot the one most significant peak SNP:

```{r}
# chr8_int_ld <- Reduce(full_join, lapply(win_chr8_int_sig_peaks$position, function(pos) {
#   return(calc_peak_ld(pos, win_chr8_int_sig, df_chr8, "allele_afr_dose_p"))
# }))
# 
# win_chr8_int_ld <- as.data.frame(inner_join(
#   win_chr8 %>% dplyr::mutate(chr = 8),
#   chr8_int_ld,
#   by = c("position" = "pos")))  # should keep all entries
# 
# chr8_int_peak <- win_chr8_int_sig_peaks$position[which.min(win_chr8_int_sig_peaks$allele_afr_dose_p)]
# 
# # Plot locus zoom
# plot_chr8_int_lz <- plot_locuszoom(win_chr8_int_ld, "8", chr8_int_peak, "allele_afr_dose_p")
# plot_chr8_int_lz

```


### **Allele Plot with Phenotype**

One for each SNP crossing significance threshold, arranged from most to least
significant. Only showing the top 6.

```{r out.width = '50%', fig.show = 'hold'}
# Inputs to plot_split should be
  # pheno_mfvc = pheno dataframe prepared above
  # "TopMed_ID" = name of ID column in pheno file
  # df_chr17 = dose frame dataframe prepared above
  # win_chr17_sig = significant LAAA results dataframe prepared above
  # .x = significant positions from win_chr17_sig prepared above
  # "maxFVC" = name of phenotype plotting - may need to update the plotting
    # function/units if changing this
  # colors = three plotting colors for manuscript
colors = c(color_lfvc, color_bfvc, color_maxfvc)

win_chr8_a_sig <- win_chr8_a_sig %>%
  dplyr::mutate(chr = "8") %>%
  dplyr::arrange(allele_dose_p)

walk(win_chr8_a_sig$position[1:6], ~ {
  print(plot_split(pheno_mfvc, "TopMed_ID", df_chr8, win_chr8_a_sig, .x, "maxFVC", colors))
})



```

### **Interaction Plot with Phenotype**

```{r out.width = '50%', fig.show = 'hold'}
# Make plots
# Inputs to plot_split should be
  # pheno_mfvc = pheno dataframe prepared above
  # "TopMed_ID" = name of ID column in pheno file
  # df_chr17 = dose frame dataframe prepared above
  # win_chr17_sig = significant LAAA results dataframe prepared above
  # .x = significant positions from win_chr17_sig prepared above
  # "maxFVC" = name of phenotype plotting - may need to update the plotting
    # function/units if changing this
  # colors = three plotting colors for manuscript
colors = c(color_lfvc, color_bfvc, color_maxfvc)

win_chr8_int_sig <- win_chr8_int_sig %>%
  dplyr::mutate(chr = "8") %>%
  dplyr::arrange(allele_afr_dose_p)

walk(win_chr8_int_sig$position[1:6], ~ {
  print(plot_split(pheno_mfvc, "TopMed_ID", df_chr8, win_chr8_int_sig, .x, "maxFVC", colors))
})

```

## **MaxFVC Chr17**

### **All Model P-Values**

```{r}
plot_sig(win_chr17)
```

Based on this, think makes sense to only look at allele dose results and
interaction results below - both anova of full model compared to null model and
African ancestry dose results are very inflated.

### **Allele & Interaction Results**

Meeting significance threshold of: `r sig_thresh`.

**All:**

Allele:

```{r}
# Get sig LAAA results & format
win_chr17_a_sig <- get_sig(win_chr17, sig_thresh, "allele_dose_p")
win_chr17_int_sig <- get_sig(win_chr17, sig_thresh, "allele_afr_dose_p")

win_chr17_a_sig %>%
  dplyr::arrange(allele_dose_p) %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)

```



Interaction:

```{r}
win_chr17_int_sig %>%
  dplyr::arrange(allele_afr_dose_p) %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)
```


### **Allele Plot with Phenotype**

One for each SNP crossing significance threshold, arranged from most to least
significant.

```{r out.width = '50%', fig.show = 'hold'}
# Inputs to plot_split should be
  # pheno_mfvc = pheno dataframe prepared above
  # "TopMed_ID" = name of ID column in pheno file
  # df_chr17 = dose frame dataframe prepared above
  # win_chr17_sig = significant LAAA results dataframe prepared above
  # .x = significant positions from win_chr17_sig prepared above
  # "maxFVC" = name of phenotype plotting - may need to update the plotting
    # function/units if changing this
  # colors = three plotting colors for manuscript
colors = c(color_lfvc, color_bfvc, color_maxfvc)

win_chr17_a_sig <- win_chr17_a_sig %>%
  dplyr::mutate(chr = "17") %>%
  dplyr::arrange(allele_dose_p)

walk(win_chr17_a_sig$position, ~ {
  print(plot_split(pheno_mfvc, "TopMed_ID", df_chr17, win_chr17_a_sig, .x, "maxFVC", colors))
})

```



## **MaxFEV1/FVC Chr10**

### **All Model P-Values**

```{r}
plot_sig(win_chr10)
```

Based on this, think makes sense to only look at allele dose results and
interaction results below - both anova of full model compared to null model and
African ancestry dose results are very inflated.

### **Allele & Interaction Results**

Meeting significance threshold of: `r sig_thresh`.

**All:**

Allele:

```{r}
# Get sig LAAA results & format
win_chr10_a_sig <- get_sig(win_chr10, sig_thresh, "allele_dose_p")
win_chr10_int_sig <- get_sig(win_chr10, sig_thresh, "allele_afr_dose_p")

win_chr10_a_sig %>%
  dplyr::arrange(allele_dose_p) %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)

```


Interaction:

```{r}
win_chr10_int_sig %>%
  dplyr::arrange(allele_afr_dose_p) %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)
```


### **Allele Plot with Phenotype**

One for each SNP crossing significance threshold, arranged from most to least
significant.

```{r out.width = '50%', fig.show = 'hold'}
colors <- c(color_lratio, color_bratio, color_maxratio)

# Inputs to plot_split should be
  # pheno_mfvc = pheno dataframe prepared above
  # "TopMed_ID" = name of ID column in pheno file
  # df_chr17 = dose frame dataframe prepared above
  # win_chr17_sig = significant LAAA results dataframe prepared above
  # .x = significant positions from win_chr17_sig prepared above
  # "maxFVC" = name of phenotype plotting - may need to update the plotting
    # function/units if changing this
  # colors = three plotting colors for manuscript
colors = c(color_lratio, color_bratio, color_maxratio)

win_chr10_a_sig <- win_chr10_a_sig %>%
  dplyr::mutate(chr = "10") %>%
  dplyr::arrange(allele_dose_p)

walk(win_chr10_a_sig$position, ~ {
  print(plot_split(pheno_mff, "TopMed_ID", df_chr10, win_chr10_a_sig, .x, "maxFEV1_FVC", colors))
})

```



# **Write Out**

Want table of significant results for each chromosome/phenotype.

```{r}
win_chr8_sig_merge_exp <- win_chr8_sig_merge %>%
  dplyr::mutate(
    chr = "8",
    pheno = "maxFVC") %>%
  dplyr::relocate(
    max_abs_beta, allele_dose_p_sig, allele_afr_dose_p_sig, .after = alt_frq) %>%
  dplyr::select(
    -n, -unadj_allele_dose_beta, -unadj_allele_dose_beta_se,
    -unadj_allele_dose_p, -allele_dose_beta_se, -afr_dose_beta_se,
    -allele_afr_dose_beta_se)

win_chr17_a_sig_exp <- win_chr17_a_sig %>%
  dplyr::mutate(
    allele_dose_p_sig = 1,
    allele_afr_dose_p_sig = 0,
    chr = "17",
    pheno = "maxFVC") %>%
  dplyr::relocate(
    max_abs_beta, allele_dose_p_sig, allele_afr_dose_p_sig, .after = alt_frq) %>%
  dplyr::select(
    -n, -unadj_allele_dose_beta, -unadj_allele_dose_beta_se,
    -unadj_allele_dose_p, -allele_dose_beta_se, -afr_dose_beta_se,
    -allele_afr_dose_beta_se)

win_chr10_a_sig_exp <- win_chr10_a_sig %>%
  dplyr::mutate(
    allele_dose_p_sig = 1,
    allele_afr_dose_p_sig = 0,
    chr = "10",
    pheno = "maxFEV1_FVC") %>%
  dplyr::relocate(
    max_abs_beta, allele_dose_p_sig, allele_afr_dose_p_sig, .after = alt_frq) %>%
  dplyr::select(
    -n, -unadj_allele_dose_beta, -unadj_allele_dose_beta_se,
    -unadj_allele_dose_p, -allele_dose_beta_se, -afr_dose_beta_se,
    -allele_afr_dose_beta_se)

write_csv(
  win_chr8_sig_merge_exp, "data/laaa_wgs/output/sig_allele_int_maxFVC_chr8.csv")
write_csv(
  win_chr17_a_sig_exp, "data/laaa_wgs/output/sig_allele_maxFVC_chr17.csv")
write_csv(
  win_chr10_a_sig_exp, "data/laaa_wgs/output/sig_allele_maxFEV1_FVC_chr10.csv")

merge_exp <- rbind(
  win_chr8_sig_merge_exp,
  win_chr17_a_sig_exp,
  win_chr10_a_sig_exp) %>%
  dplyr::relocate(pheno, chr, .before = position)

write_csv(
  merge_exp, "data/laaa_wgs/output/sig_all_results_all_chr.csv")

```


# **Session Info**

```{r session_info}
sessionInfo()
```
