---
title: "Local ancestry inference QC"
author: "SDS"
date: "Last compiled on `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: kable
    code_folding: hide
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = paste0(Sys.getenv("RKJCOLLAB"), "/Collabs/ortega/reports")) })
---

```{r setup, include = F}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(
  root.dir = paste0(Sys.getenv("RKJCOLLAB"), "/Collabs/ortega"))

library(tidyverse)
library(conflicted)

```

# **Overview**

**Current:** Script now focuses on global ancestry calculated from RFMix and
ADMIXTURE on the unimputed data run by both Meher and SDS. For final analyses,
will be using everything calculated by SDS so that can accurately report all
code and software versions used. Focus on unimputed data because the
RFMix-calculated global ancestry estimate using unimputed data is used as a
covariate for both admixture mapping and LAAA.

*TODO: may add updated QC once re-run of RFMix on WGS data is ready.* 

**Previous:** This script originally looked at the quality of RFMIx results run
by SDS on SARP-CSGA unimputed and imputed as well as WGS RFMix results for chr8
only run by Meher. SDS did not re-run WGS RFMix because missing TOPMed freeze 8
chr2 on server. Previous version was before conversation with Michelle that
revealed the issue with RFMix was due to not correctly extracting the phasing
information available in WGS and TOPMed-imputed data.

# **Load Data**

Unimputed results were calculated by both Meher and SDS, from PLINK files at:
Victor_Ortega_analyses_MPB/SARP_CSGA_merge/SARP123_CSGA_merged_final. One
version of both RFMix and ADMIXTURE global ancestry estimates were generated by
Meher, and one version of RFMix and two versions of ADMIXTURE global ancestry
estimates were generated by SDS. SDS originally generated a version of ADMIXTURE
results that matched Meher's results, but noted warnings about duplicate
position SNPs with different variant IDs not merging between the study data and
reference data in PLINK before ADMIXTURE. After fixing those warnings (by
matching ref/alt alleles in study data to reference data before merge where
possible), SDS results differ from Meher's results, but more closely match the
RFMix-calculated global ancestry estimates.

Imputed results used unimputed input above, and had two different filtering
criteria applied following imputation:

  + original: MAF 0.01 & RSQ 0.3
  + strict: MAF 0.03 & RSQ 0.9
  
WGS RFMix results are from:
Victor_Ortega_analyses_MPB/SARP_TOPMED_WGS_data/rfmix_wgs/chr8.

```{r, message = F}
# Load merged gwide from all methods
gwide_imp <- read.delim(
  "data/gwide_anc_imp/maf0.01_rsq0.3/merged_gwide.txt")
gwide_imp_strict <- read.delim(
  "data/gwide_anc_imp/maf0.03_rsq0.9/merged_gwide.txt")
gwide_unimp <- read.delim(
  "data/gwide_anc_unimp/merged_gwide.txt")
gwide_unimp_prev <- read.delim(
  "data/gwide_anc_unimp/pre_merge_fix/merged_gwide.txt")
gwide_unimp_meher <-
  read.delim(paste0(
    "background/from_dayam_server/Victor_Ortega_analyses_MPB/SARP_CSGA_merge/",
    "rfmix/data/output/merged_gwide.txt"))

# Load WGS RFMix gwide only since didn't have all chr for ADMIXTURE
rfmix_gwide_wgs <- read_delim("data/rfmix_wgs/rfmix_gwide.txt")


# Load local ancestry segments from all methods
local_anc_imp <- read.delim(
  "data/gwide_anc_imp/maf0.01_rsq0.3/local_ancestry_segments.txt")
local_anc_imp_strict <- read.delim(
  "data/gwide_anc_imp/maf0.03_rsq0.9/local_ancestry_segments.txt")
local_anc_unimp <- read.delim(
  "data/gwide_anc_unimp/local_ancestry_segments.txt")
local_anc_unimp_prev <- read.delim(
  "data/gwide_anc_unimp/pre_merge_fix/local_ancestry_segments.txt")
local_anc_unimp_meher <-
  read.delim(paste0(
    "background/from_dayam_server/Victor_Ortega_analyses_MPB/SARP_CSGA_merge/",
    "rfmix/data/output/local_ancestry_segments.txt"))

# Add RFMix directories from all methods
rfmix_dir_imp <- "data/rfmix_imp/output_co/maf0.01_rsq0.3"
rfmix_dir_imp_strict <- "data/rfmix_imp/output_co/maf0.03_rsq0.9"
rfmix_dir_unimp <- "data/rfmix_unimp/output_co"

  
```

# **Current Version**

## **Genome-wide ancestry differences**

```{r}
# Function for this section
show_gwide_anc_diff <- function(df) {
delta_unimp <- df$RFMIX_GW_AFR - df$ADMIXTURE_GW_AFR
print(summary(delta_unimp))
hist(delta_unimp, breaks=50, main="Genome-wide ancestry estimate differences")

ggplot(df, aes(RFMIX_GW_AFR, ADMIXTURE_GW_AFR)) +
  geom_point() +
  xlim(c(0,1)) +
  ylim(c(0,1)) +
  coord_fixed(ratio = 1)
}
```


Delta = RFMIX estimate - ADMIXTURE estimate

### SARP-CSGA Unimputed with Merge Fix

RFMix run by Sarah, with code updated before ADMIXTURE to fix merge issue with
reference (described above in load data section).

```{r, out.width = '50%', fig.show = 'hold'}
show_gwide_anc_diff(gwide_unimp)
```

Really great concordance between estimates calculated with ADMIXTURE and those
calculated by RFMix.

### SARP-CSGA Unimputed

RFMix run by Sarah, with same code as Meher.

```{r, out.width = '50%', fig.show = 'hold'}
show_gwide_anc_diff(gwide_unimp_prev)
```

Calculated from RFMix skewed higher than ADMIXTURE estimates.

### SARP-CSGA Unimputed by Meher

RFMix run by Meher.

```{r, out.width = '50%', fig.show = 'hold'}
show_gwide_anc_diff(gwide_unimp_meher)
```

Calculated from RFMix also skewed higher than ADMIXTURE estimates.

### Compare Unimputed by Meher & Sarah

**ADMIXTURE Estimates:**

```{r out.width = '33%', fig.show = 'hold'}
gwide_unimp_meher_prev <- inner_join(gwide_unimp_prev, gwide_unimp_meher, by = c("IID"),
                                suffix = c("_sds_prev", "_meher"))
gwide_unimp_meher_merge_fix <- inner_join(gwide_unimp, gwide_unimp_meher, by = c("IID"),
                                suffix = c("_sds_merge_fix", "_meher"))
gwide_unimp_prev_merge_fix <- inner_join(gwide_unimp_prev, gwide_unimp, by = c("IID"),
                                suffix = c("_sds_prev", "_sds_merge_fix"))

ggplot(gwide_unimp_meher_prev, aes(ADMIXTURE_GW_AFR_sds_prev, ADMIXTURE_GW_AFR_meher)) +
  geom_point() +
  xlim(c(0,1)) +
  ylim(c(0,1)) +
  coord_fixed(ratio = 1)

ggplot(gwide_unimp_meher_merge_fix, aes(ADMIXTURE_GW_AFR_sds_merge_fix, ADMIXTURE_GW_AFR_meher)) +
  geom_point() +
  xlim(c(0,1)) +
  ylim(c(0,1)) +
  coord_fixed(ratio = 1)

ggplot(gwide_unimp_prev_merge_fix, aes(ADMIXTURE_GW_AFR_sds_prev, ADMIXTURE_GW_AFR_sds_merge_fix)) +
  geom_point() +
  xlim(c(0,1)) +
  ylim(c(0,1)) +
  coord_fixed(ratio = 1)
```

SDS version before merge fix matches Meher's version. SDS version after merge
fix has higher AFR ancestry.

**RFMix Estimates:**

```{r out.width = '33%', fig.show = 'hold'}
ggplot(gwide_unimp_meher_prev, aes(RFMIX_GW_AFR_sds_prev, RFMIX_GW_AFR_meher)) +
  geom_point() +
  xlim(c(0,1)) +
  ylim(c(0,1)) +
  coord_fixed(ratio = 1)

ggplot(gwide_unimp_meher_merge_fix, aes(RFMIX_GW_AFR_sds_merge_fix, RFMIX_GW_AFR_meher)) +
  geom_point() +
  xlim(c(0,1)) +
  ylim(c(0,1)) +
  coord_fixed(ratio = 1)

ggplot(gwide_unimp_prev_merge_fix, aes(RFMIX_GW_AFR_sds_prev, RFMIX_GW_AFR_sds_merge_fix)) +
  geom_point() +
  xlim(c(0,1)) +
  ylim(c(0,1)) +
  coord_fixed(ratio = 1)

```

All RFMix global ancestry estimates are concordant.

**Conclusion:** Even though SDS updated version with code fixed for better
merge does not match Meher's results, it does have better concordance with the
RFMix-calculated global ancestry estimates. Think should use ADMIXTURE results
from the merge fixed code. Since the RFMix-calculated estimates are what are
actually used as covariates, this likely won't affect any results, but want to
clarify the ADMIXTURE version in the stable results here.

## **Local ancestry deviation distribution**

### SARP-CSGA Unimputed

```{r, echo=FALSE}
mean_afr_anc_unimp <- mean(gwide_unimp$RFMIX_GW_AFR)
dev_unimp <- local_anc_unimp$prop.afr - mean_afr_anc_unimp
z_unimp <- dev_unimp/sd(dev_unimp)
hist(z_unimp, breaks=100, main="Histogram of local ancestry deviations",
     xlim = c(-4, 3))
```


## **Local ancestry across the genome**

### SARP-CSGA Unimputed

```{r, echo=FALSE, fig.height=12, fig.width=9}
draw.frame <- data.frame()
for (chr in 1:22) {
  chr.anc <- local_anc_unimp[local_anc_unimp$chr.nr == chr,] 
  chr.anc$end.snp.nr <- cumsum(chr.anc$nr.snps)
  chr.anc$start.snp.nr <- c(1, chr.anc$end.snp.nr[-length(chr.anc$end.snp.nr)] + 1)
  snps <- read.table(
    paste0(rfmix_dir_unimp, "/chr",
           chr, "/chr", chr, "_local_ancestry_snps.txt"))[,1]
  chr.anc$begin.pos <- snps[chr.anc$start.snp.nr]
  chr.anc$end.pos <- snps[chr.anc$end.snp.nr]
  chr.anc <- chr.anc[order(chr.anc$seg.nr),]
  chr.anc$pos <- rowMeans(chr.anc[,c(dim(chr.anc)[2]-1,dim(chr.anc)[2])])
  draw.frame <- rbind(draw.frame,
                      data.frame(Chromosome=paste("Chromosome", rep(chr, dim(chr.anc)[1])),
                                 Position=chr.anc$pos/10000000,
                                 Ancestry=chr.anc$prop.afr.anc))
}
draw.frame$Chromosome <- factor(draw.frame$Chromosome, 
                                levels=c(
                                  "Chromosome 1", "Chromosome 2", "Chromosome 3", "Chromosome 4",
                                  "Chromosome 5", "Chromosome 6", "Chromosome 7", "Chromosome 8",
                                  "Chromosome 9","Chromosome 10","Chromosome 11","Chromosome 12",
                                  "Chromosome 13","Chromosome 14","Chromosome 15","Chromosome 16",
                                  "Chromosome 17","Chromosome 18","Chromosome 19","Chromosome 20",
                                  "Chromosome 21","Chromosome 22"))

ggplot(draw.frame, aes(Position, Ancestry)) +  
  facet_wrap( ~ Chromosome, nrow=6, ncol=4) + 
  geom_line(linewidth=0.2) + 
  labs(x = expression(paste("Base pair position ", (10^7)))) + 
  labs(y = "Proportion African ancestry") +
  theme_bw() +
  theme(legend.position="top") + 
  scale_color_brewer(palette="Set1") +
  geom_hline(aes(yintercept=mean_afr_anc_unimp), size=0.0725)
```

# **Previous Version**

## **Genome-wide ancestry differences**

### SARP-CSGA WGS

Like described at top of report, only have chr8 from RFMix run by Meher, so
only showing distribution of the global ancestry estimates as calculated by
chr8.

**Note the scale of the following plot:**

```{r}
ggplot(rfmix_gwide_wgs, aes(RFMIX_GW_AFR)) +
  geom_histogram()
```


### SARP-CSGA Imputed Original

Below is a summary and histogram of the differences between genome-wide ancestry
estimated using ADMIXTURE, and genome-wide ancestry estimated using RFMix, per
individual.

```{r, out.width = '50%', fig.show = 'hold'}
delta_imp <- gwide_imp$RFMIX_GW_AFR - gwide_imp$ADMIXTURE_GW_AFR
summary(delta_imp)
hist(delta_imp, breaks=50, main="Genome-wide ancestry estimate differences")

ggplot(gwide_imp, aes(RFMIX_GW_AFR, ADMIXTURE_GW_AFR)) +
  geom_point() +
  xlim(c(0,1)) +
  ylim(c(0,1)) +
  coord_fixed(ratio = 1)
```

This looks really bad... Clearly something wrong with RFMix estimates since
would not expect this population to be so close to 1.0 YRI.

### SARP-CSGA Imputed Strict

After some research, determined that low quality imputation results could be 
affecting RFMix estimates. Based on Michelle's paper
(PMID 34762840)[https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8787857/], using
MAF 0.03 and RSQ 0.9 after imputation.


```{r, out.width = '50%', fig.show = 'hold'}
delta_imp_strict <- gwide_imp_strict$RFMIX_GW_AFR - gwide_imp_strict$ADMIXTURE_GW_AFR
summary(delta_imp_strict)
hist(delta_imp_strict, breaks=50, main="Genome-wide ancestry estimate differences")

ggplot(gwide_imp_strict, aes(RFMIX_GW_AFR, ADMIXTURE_GW_AFR)) +
  geom_point() +
  xlim(c(0,1)) +
  ylim(c(0,1)) +
  coord_fixed(ratio = 1)
```

### Compare Imputed Original and Strict

```{r out.width = '50%', fig.show = 'hold'}
gwide_imp_comb <- inner_join(gwide_imp, gwide_imp_strict,
                             by = c("IID"), suffix = c("_orig", "_strict"))
ggplot(gwide_imp_comb, aes(RFMIX_GW_AFR_orig, RFMIX_GW_AFR_strict)) +
  geom_point() +
  xlim(c(0,1)) +
  ylim(c(0,1)) +
  coord_fixed(ratio = 1)

ggplot(gwide_imp_comb, aes(ADMIXTURE_GW_AFR_orig, ADMIXTURE_GW_AFR_strict)) +
  geom_point() +
  xlim(c(0,1)) +
  ylim(c(0,1)) +
  coord_fixed(ratio = 1)
```

There doesn't appear to be a significant difference between these.


## **Local ancestry deviation distribution**

### SARP-CSGA Imputed Original

The deviation between mean local African ancestry and mean genome-wide African
ancestry was calculated for each local ancestry segment. 

```{r}
mean_afr_anc_imp <- mean(gwide_imp$RFMIX_GW_AFR)
dev <- local_anc_imp$prop.afr - mean_afr_anc_imp
z <- dev/sd(dev)
hist(z, breaks=100, main="Histogram of local ancestry deviations",
     xlim = c(-4, 3))
```

### SARP-CSGA Imputed Strict

```{r}
mean_afr_anc_imp_strict <- mean(gwide_imp_strict$RFMIX_GW_AFR)
dev <- local_anc_imp_strict$prop.afr - mean_afr_anc_imp_strict
z <- dev/sd(dev)
hist(z, breaks=100, main="Histogram of local ancestry deviations",
     xlim = c(-4, 3))
```


## **Local ancestry across the genome**

### SARP-CSGA Imputed Original

Below is a plot of the mean local ancestry across the genome.

```{r, echo=FALSE, fig.height=12, fig.width=9}
draw.frame <- data.frame()
for (chr in 1:22) {
  chr.anc <- local_anc_imp[local_anc_imp$chr.nr == chr,] 
  chr.anc$end.snp.nr <- cumsum(chr.anc$nr.snps)
  chr.anc$start.snp.nr <- c(1, chr.anc$end.snp.nr[-length(chr.anc$end.snp.nr)] + 1)
  snps <- read.table(
    paste0(rfmix_dir_imp, "/chr",
           chr, "/chr", chr, "_local_ancestry_snps.txt"))[,1]
  chr.anc$begin.pos <- snps[chr.anc$start.snp.nr]
  chr.anc$end.pos <- snps[chr.anc$end.snp.nr]
  chr.anc <- chr.anc[order(chr.anc$seg.nr),]
  chr.anc$pos <- rowMeans(chr.anc[,c(dim(chr.anc)[2]-1,dim(chr.anc)[2])])
  draw.frame <- rbind(draw.frame,
                      data.frame(Chromosome=paste("Chromosome", rep(chr, dim(chr.anc)[1])),
                                 Position=chr.anc$pos/10000000,
                                 Ancestry=chr.anc$prop.afr.anc))
}
draw.frame$Chromosome <- factor(draw.frame$Chromosome, 
                                levels=c(
                                  "Chromosome 1", "Chromosome 2", "Chromosome 3", "Chromosome 4",
                                  "Chromosome 5", "Chromosome 6", "Chromosome 7", "Chromosome 8",
                                  "Chromosome 9","Chromosome 10","Chromosome 11","Chromosome 12",
                                  "Chromosome 13","Chromosome 14","Chromosome 15","Chromosome 16",
                                  "Chromosome 17","Chromosome 18","Chromosome 19","Chromosome 20",
                                  "Chromosome 21","Chromosome 22"))

ggplot(draw.frame, aes(Position, Ancestry)) +  
  facet_wrap( ~ Chromosome, nrow=6, ncol=4) + 
  geom_line(linewidth=0.2) + 
  labs(x = expression(paste("Base pair position ", (10^7)))) + 
  labs(y = "Proportion African ancestry") +
  theme_bw() +
  theme(legend.position="top") + 
  scale_color_brewer(palette="Set1") +
  geom_hline(aes(yintercept=mean_afr_anc_imp), size=0.0725)
```

### SARP-CSGA Imputed Strict

```{r, echo=FALSE, fig.height=12, fig.width=9}
draw.frame <- data.frame()
for (chr in 1:22) {
  chr.anc <- local_anc_imp_strict[local_anc_imp_strict$chr.nr == chr,] 
  chr.anc$end.snp.nr <- cumsum(chr.anc$nr.snps)
  chr.anc$start.snp.nr <- c(1, chr.anc$end.snp.nr[-length(chr.anc$end.snp.nr)] + 1)
  snps <- read.table(
    paste0(rfmix_dir_imp_strict, "/chr",
           chr, "/chr", chr, "_local_ancestry_snps.txt"))[,1]
  chr.anc$begin.pos <- snps[chr.anc$start.snp.nr]
  chr.anc$end.pos <- snps[chr.anc$end.snp.nr]
  chr.anc <- chr.anc[order(chr.anc$seg.nr),]
  chr.anc$pos <- rowMeans(chr.anc[,c(dim(chr.anc)[2]-1,dim(chr.anc)[2])])
  draw.frame <- rbind(draw.frame,
                      data.frame(Chromosome=paste("Chromosome", rep(chr, dim(chr.anc)[1])),
                                 Position=chr.anc$pos/10000000,
                                 Ancestry=chr.anc$prop.afr.anc))
}
draw.frame$Chromosome <- factor(draw.frame$Chromosome, 
                                levels=c(
                                  "Chromosome 1", "Chromosome 2", "Chromosome 3", "Chromosome 4",
                                  "Chromosome 5", "Chromosome 6", "Chromosome 7", "Chromosome 8",
                                  "Chromosome 9","Chromosome 10","Chromosome 11","Chromosome 12",
                                  "Chromosome 13","Chromosome 14","Chromosome 15","Chromosome 16",
                                  "Chromosome 17","Chromosome 18","Chromosome 19","Chromosome 20",
                                  "Chromosome 21","Chromosome 22"))

ggplot(draw.frame, aes(Position, Ancestry)) +  
  facet_wrap( ~ Chromosome, nrow=6, ncol=4) + 
  geom_line(linewidth=0.2) + 
  labs(x = expression(paste("Base pair position ", (10^7)))) + 
  labs(y = "Proportion African ancestry") +
  theme_bw() +
  theme(legend.position="top") + 
  scale_color_brewer(palette="Set1") +
  geom_hline(aes(yintercept=mean_afr_anc_imp_strict), size=0.0725)
```


# **Session Info**

```{r}
sessionInfo()
```
