---
title: "7.4_laaa_results"
author: "SDS"
date: "Last compiled on `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: kable
    code_folding: hide
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = paste0(Sys.getenv("RKJCOLLAB"), "/Collabs/ortega/reports")) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(
  root.dir = paste0(Sys.getenv("RKJCOLLAB"), "/Collabs/ortega"))

# Common packages
library(tidyverse)
library(knitr)
library(biomaRt)
library(DT)
library(conflicted)

#  For locuszoomr
library(locuszoomr)
library(ensembldb)
library(AnnotationHub)
library(GGally)
library(RColorBrewer)
library(ggpubr)
library(IRanges)
library(snpStats)

# Set display options
options(digits = 4)  # only 2 digits after decimal

```

# **Overview**

Results of LAAA run by Sarah on local ancestry results from WGS data (RFMix
run by Sarah) for admixture mapping peaks found on chromosomes 8, 10, and 17
for maxFVC and maxFEV1/FVC (see report "15_summary_rerun_admix_mapping").

Instead of showing results exactly as Michelle did in 
[PMID 34762840](https://pubmed.ncbi.nlm.nih.gov/34762840/), going to evaluate
as suggested in original LAAA paper
[PMID 29226381](https://www.ncbi.nlm.nih.gov/pmc/articles/pmid/29226381/).
Michelle checked for significance of p-allele < 10^-3 (0.001). The LAAA
paper suggests first checking for overall model significance at a specific
variant, then determining which part of the model is driving that significance
by "selecting the effect with the largest absolute value of the test
statistics".

Interpretation of overall model significance is that when fine map, the LAAA
model significantly explains variation in maxFVC at the a. priori selected
threshold of model P < 10^-3.


**LAAA model formula:**

pheno ~ group + age + sex + ht + bmi + RFMIX_GW_AFR + allele_dose + afr_dose + allele_afr_dose

Group is SARP12+CSGA & SARP3, and RFMIX_GW_AFR is global African ancestry
proportion calculated from unimputed chip data (same as admixture mapping input).

*Note that admixture mapping results shown in h19, but LAAA/fine mapping*
*results shown in hg38.*

For locus zoom plots, LD is calculated within our study population using
R package snpStats.

# **Load Data**

```{r load_data, message = F}

# LAAA results - hg38
win_chr5 <- read_delim(
  "data/laaa_wgs_new_2/output/region_hg38_maxFVC_chr5/maxFVC_assoc_chr5.txt")
win_chr8 <- read_delim(
  "data/laaa_wgs_new_2/output/region_hg38_maxFVC_chr8/maxFVC_assoc_chr8.txt")
win_chr10 <- read_delim(
  "data/laaa_wgs_new_2/output/region_hg38_maxFEV1_FVC_chr10/maxFEV1_FVC_assoc_chr10.txt")
win_chr17 <- read_delim(
  "data/laaa_wgs_new_2/output/region_hg38_maxFVC_chr17/maxFVC_assoc_chr17.txt")

# Dose frames
read_dose_frames <- function(dir_name) {
    df_list <- list.files(dir_name, full.names = T)
    
    df_all <- Reduce(rbind, lapply(df_list, function(file) {
      file_base <- basename(file)
      df_name <- gsub(
        ".+[[:digit:]]+_", "",
        gsub("\\.txt", "", file_base))
      df <- read_delim(file) %>%
        dplyr::mutate(type = df_name) %>%
        relocate(type)
      return(df)
  }))
  return(df_all)
}

df_chr5 <-
  read_dose_frames("data/laaa_wgs_new_2/dose_frames/region_hg38_maxFVC_chr5")
df_chr8 <-
  read_dose_frames("data/laaa_wgs_new_2/dose_frames/region_hg38_maxFVC_chr8")
df_chr10 <-
  read_dose_frames("data/laaa_wgs_new_2/dose_frames/region_hg38_maxFEV1_FVC_chr10")
df_chr17 <-
  read_dose_frames("data/laaa_wgs_new_2/dose_frames/region_hg38_maxFVC_chr17")

# Pheno values
pheno_mff <- read_delim(
  "data/pheno/SARP123_CSGA_348_laaa_maxFEV1_FVC_pheno.txt")
pheno_mfvc <- read_delim(
  "data/pheno/SARP123_CSGA_348_laaa_maxFVC_pheno.txt")

# Set main paper colors
source("/Users/slacksa/repos/pgx_laba/set_main_paper_colors.R")

```

**Regions Tested:**

+ MaxFVC chr5:`r format(min(win_chr5$position), scientific = F)` - `r format(max(win_chr5$position), scientific = F)`
+ MaxFVC chr8:`r format(min(win_chr8$position), scientific = F)` - `r format(max(win_chr8$position), scientific = F)`
+ MaxFVC chr17:`r format(min(win_chr17$position), scientific = F)` - `r format(max(win_chr17$position), scientific = F)`
+ MaxFEV1/FVC chr10:`r format(min(win_chr10$position), scientific = F)` - `r format(max(win_chr10$position), scientific = F)`

# **Functions**

**For Report:**

```{r}
# Significance thresholds
sig_thresh <- 0.001
nom_sig_thresh <- 0.01

# Colors
# colors <- brewer.pal(3, "BuGn")  # used for publication colors made

# Get significant results
get_sig <- function(df, thresh) {
  # Get ANOVA significant results
  df_sig <- df %>% dplyr::filter(anova_p < thresh)
  
  # Write abs(max) column name in new column
  df_sig$max_abs_beta <- apply(
    df_sig[, c("allele_dose_beta", "afr_dose_beta", "allele_afr_dose_beta")],
    1, function(row) {
      cols <- c("allele_dose_beta", "afr_dose_beta", "allele_afr_dose_beta")
      cols[which.max(abs(row))]
  })
  return(df_sig)
}

# Make plotting function
plot_sig <- function(win_df) {

  win_df_lg <- win_df %>%
    dplyr::select(position, allele_dose_p, afr_dose_p, allele_afr_dose_p, anova_p) %>%
    pivot_longer(cols = c(allele_dose_p, afr_dose_p, allele_afr_dose_p, anova_p),
                 names_to = "type",
                 values_to = "pval")
  
  win_df_nom_sig <- subset(win_df_lg, pval < nom_sig_thresh & type == "anova_p")
  
  ggplot(win_df_lg, aes(position, -log10(pval))) +
    geom_point() +
    geom_point(data = subset(win_df_lg, position %in% win_df_nom_sig$position),
               aes(color = "orange")) +
    labs(x = "Position", y = "-log10(P-Value)", color = "significant") +
    facet_wrap(~ type) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

fun_split <- function(x, data, pheno) {
  return(data.frame(
    y = min(data[[pheno]]) - (0.1*(max(data[[pheno]]) - min(data[[pheno]]))),
    label = length(x)))  # x comes from group = Asthma
}

# Plot split phenotype


pheno_data <- pheno_mfvc
dose_frame <- df_chr8
sig_data <- win_chr8_sig
position <- sig_pos
pheno <- "maxFVC"
colors <- colors

plot_split <- function(pheno_data, dose_frame, sig_data, position, pheno, colors) {
 
  # Merge pheno and dose frame
  df_afr <- dose_frame %>% dplyr::filter(type == "afr_dose") %>%
    dplyr::select(-type, -ref, -alt) %>%
    column_to_rownames(var = "position")
  df_afr_t <- as.data.frame(t(df_afr))
  colnames(df_afr_t) <- paste0(colnames(df_afr_t), "_afr")
  df_afr_t <- rownames_to_column(df_afr_t, var = "ID")
  
  df_allele <- dose_frame %>% dplyr::filter(type == "allele_dose") %>%
    dplyr::select(-type, -ref, -alt) %>%
    column_to_rownames(var = "position")
  df_allele_t <- as.data.frame(t(df_allele))
  colnames(df_allele_t) <- paste0(colnames(df_allele_t), "_allele")
  df_allele_t <- rownames_to_column(df_allele_t, var = "ID")
  
  pheno_df_temp <- inner_join(
    pheno_data,
    df_afr_t,
    by = c("TopMed_ID" = "ID"))
  pheno_df <- inner_join(
    pheno_df_temp,
    df_allele_t,
    by = c("TopMed_ID" = "ID"))
  
  # Get additional details for plot from sig_data
  driving_term <- sig_data$max_abs_beta[sig_data$position == position]
  
  # Make format of allele dose that have alt & ref alleles
  alt <- sig_data$alt[sig_data$position == position]
  ref <- sig_data$ref[sig_data$position == position]
  allele_labels <- c(
    `0` = paste0(ref, ref),
    `1` = paste0(ref, alt),
    `2` = paste0(alt, alt))
  
  plot <- ggplot(pheno_df, aes(
    x = factor(!!rlang::sym(paste0(position, "_allele"))),
    y = !!rlang::sym(pheno))) +
  geom_boxplot(aes(fill = "Combined"),
               varwidth = TRUE, alpha = 0.7, outlier.shape = 21) +
  geom_boxplot(aes(fill = factor(!!rlang::sym(paste0(position, "_afr")))),
               varwidth = TRUE, alpha = 0.7, outlier.shape = 21) +
  labs(x = "Genotype",
       y = paste0(
         gsub("_", "/", gsub("m", "M", pheno)),
         ifelse(str_detect(pheno, "_"), " Ratio", " (mL)")),
       fill = "Copies Local African Ancestry",
       title = paste0(
         "chr", sig_data$chr, ":", position, ":", ref, ":", alt, ", Driving Term: ", driving_term)) +
  scale_x_discrete(labels = allele_labels) +
  scale_fill_manual(values = c("Combined" = "transparent",  # Transparent box plot
                             "0" = colors[1], "1" = colors[2], "2" = colors[3]),
                    labels = c("Combined" = "Combined", "0" = "0",
                               "1" = "1", "2" = "2")) +
  theme_bw() +
  theme(legend.position = "bottom") +
  stat_summary(fun.data = fun_split, fun.args = list(data = pheno_df, pheno = pheno),
               geom = "text", aes(group = !!rlang::sym(paste0(position, "_afr"))),
               position = position_dodge(0.5), hjust = 0.5)
  return(plot)
}

# Get LD
calc_peak_ld <- function(peak_snp, win_sig, dose_frame) {
  # Get peak snp
  peak <- paste0(
    win_sig$position[which.min(win_sig$anova_p)], "_allele")
  
  # Modify dose frame, use to create SnpMatrix
  df_allele_t <- as.data.frame(t(dose_frame %>%
    dplyr::filter(type == "allele_dose") %>%
    dplyr::select(-type, -ref, -alt) %>%
    column_to_rownames(var = "position")))
  snp_m <- new("SnpMatrix", as.matrix(df_allele_t))
  
  # Get pairwise LD with peak SNP
  ld <- ld(snp_m[, which(colnames(snp_m) == peak_snp), drop = F],
           snp_m, stats = c("R.squared"))
  
  # LD R2 column name is just position of peak SNP calculated with respect to
  ld_t <- as.data.frame(t(ld)) %>%
    rownames_to_column("pos") %>%
    dplyr::mutate(pos = as.integer(gsub("_allele", "", pos)))
  
  return(ld_t)
}

# Plot locus zoom
plot_locuszoom <- function(win_ld, chr, peak_snp) {
  loc <- locus(data = win_ld, chrom = "chr", p = "anova_p", pos = "position",
             ens_db = ensDb_v111, seqname = chr, labs = "position",
             xrange = c(min(win_ld$position), max(win_ld$position)),
             LD = as.character(peak_snp))
  loc_plot <- locus_ggplot(loc, filter_gene_biotype = "protein_coding",
             pcutoff = nom_sig_thresh)
  return(loc_plot)
}

```


# **Results**

For each window, first going to check for overall model significance. For each
variant that is significant, then going to identify the effect with the largest
absolute value of the test statistics to determine which part of the LAAA model
is driving that significance.

For significance threshold, going to follow Dr. Ortega's and Michelle's previous
paper: P < 10^-3 (0.001).

For nominal significance, going to look at results with model P-value < 0.01.

## **MaxFVC Chr5**

### **LAAA Results**

Looking at all results meeting at least nominal significance: `r nom_sig_thresh`

```{r }
# Get sig LAAA results & format
win_chr5_sig <- get_sig(win_chr5, nom_sig_thresh)

win_chr5_sig <- win_chr5_sig %>%
  dplyr::rowwise() %>%
  dplyr::mutate(chr = 5) %>%
  dplyr::relocate(anova_p, max_abs_beta) %>%
  dplyr::arrange(anova_p)

win_chr5_sig %>%
  dplyr::select(
    anova_p, max_abs_beta, position, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)
# datatable(win_chr5_sig)

```

No significant results - same as before applied MAF filter ahead of RFMix.

### **Plot Model P-Values**

```{r, warning = F}
plot_sig(win_chr5)
```


## **MaxFVC Chr8**

### **LAAA Results**

Looking at all results meeting at least nominal significance: `r nom_sig_thresh`.

**All:**

```{r }
# Get sig LAAA results & format
win_chr8_sig <- get_sig(win_chr8, nom_sig_thresh)

win_chr8_sig <- win_chr8_sig %>%
  dplyr::rowwise() %>%
  dplyr::mutate(chr = 8) %>%
  dplyr::relocate(anova_p, max_abs_beta) %>%
  dplyr::arrange(anova_p)

win_chr8_sig %>%
  dplyr::select(
    position, anova_p, max_abs_beta, position, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)
# datatable(win_chr8_sig)

```

Results include 3 out of 4 of the results before MAF filter was correctly
applied before RFMix, and many additional new results.

Confirmed the 1 result no longer significant was still in input, just that
anova_p no longer meets significance threshold.

**Summarized to keep only the first of those in complete LD:**

For now, doing this by filtering to keep only one of multiple rows with
identical alt_frq, anova_p, and max_abs_beta.

```{r}
win_chr8_sig_dedup <- win_chr8_sig %>%
  distinct(alt_frq, anova_p, max_abs_beta, .keep_all = TRUE)

win_chr8_sig_dedup %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)
```

Still have 16 results! How best to dig into these? Almost all are driven by
the allele_afr_dose_beta. May help to see if any map to genes.

For locus Zoom plot, want one version for each of the "peaks". To do this,
need to narrow down the above to keep only top SNP from each set of physically
close SNPs.

```{r}
# Get ranges of points within 30,000bp of each other
win_chr8_sig_ranges <-
  IRanges(start = win_chr8_sig$position, end = win_chr8_sig$position)
win_chr8_sig_merged_ranges <-
  IRanges::reduce(win_chr8_sig_ranges, min.gapwidth = 30000)

# Add range group to original df
win_chr8_sig$group <-
  subjectHits(findOverlaps(win_chr8_sig_ranges, win_chr8_sig_merged_ranges))

# Now, modify df to keep only most significant pval in each group

# Group by the merged ranges and keep the row with the lowest p-value in each group
win_chr8_sig_peaks <- win_chr8_sig %>%
  dplyr::group_by(group) %>%
  dplyr::slice(which.min(anova_p)) %>%
  ungroup() %>%
  arrange(anova_p)

head(win_chr8_sig_peaks)
```


### **Locus Zoom Plot**

Only plot most significant peak SNP:

```{r}
# Get genome annotation
ah <- AnnotationHub()
# query(ah, c("EnsDb", "Homo sapiens"))
ensDb_v111 <- ah[["AH116291"]]

chr8_ld <- Reduce(full_join, lapply(win_chr8_sig_peaks$position, function(pos) {
  return(calc_peak_ld(pos, win_chr8_sig, df_chr8))
}))

win_chr8_ld <- as.data.frame(inner_join(
  win_chr8 %>% dplyr::mutate(chr = 8),
  chr8_ld,
  by = c("position" = "pos")))  # should keep all entries


chr8_peak <- win_chr8_sig_peaks$position[which.min(win_chr8_sig_peaks$anova_p)]

# Plot locus zoom
plot_chr8_lz <- plot_locuszoom(win_chr8_ld, "8", chr8_peak)

```


### **Plot with Phenotype**

One for each SNP crossing our significance threshold, arranged from most to
least significant.

```{r out.width = '50%', fig.show = 'hold'}
# Make plots
# Using walk to suppress ## 1, 2, output during lapply
# walk(win_chr8_sig_peaks$position, ~ {
#   print(plot_split(pheno_mfvc, df_chr8, win_chr8_sig, .x, "maxFVC"))
# })

```

**Example for position 55541918 updated compared to previous results:**

Updated TT category has 10 with 0 copies of LAA, 21 with 1 copy, 4 with 2.
Previous TT category had 8 with 0 copies of LAA, 23 with 1 copy, 4 with 2.

Updated TC category has 93 with 0 copies, 82 with 1 copy, 4 with 2.
Previous TC category had 92 with 0 copies, 85 with 1 copy, 2 with 2.

Updated CC category has 65 with 0 copies, 64 with 1, 5 with 2.
Previous CC category had 65 with 0 copies, 64 with 1, 5 with 2.


### **BiomaRt Lookup**

```{r}
mart <- useMart("ensembl", dataset="hsapiens_gene_ensembl")  # default is hg38

win_chr8_sig$chr = 8
result <- NA

result <- do.call(rbind, lapply(1:nrow(win_chr8_sig), function(i) {
  result <- getBM(
    attributes = c('ensembl_gene_id', 'hgnc_symbol', 'chromosome_name', 'start_position', 'end_position'),
    filters = c('chromosome_name','start','end'),
    values = list(win_chr8_sig$chr[i], win_chr8_sig$position[i], win_chr8_sig$position[i]),
    mart = mart)
}))

head(result)

```

All the SNPs combined contain genes:

+ ENSG00000206579 XKR4
+ ENSG00000254254 PENK-AS1
+ ENSG00000246430 LINC00968 


**More about XKR4:**

From the human protein atlas: transporter membrane protein. Phospholipid
scramblase that promotes phosphatidylserine exposure on apoptotic cell surface.
Enables phospholipid scramblase activity. Involved in phosphatidylserine
exposure on apoptotic cell surface. Predicted to be integral component of
membrane. Predicted to be active in plasma membrane.

Looks like has been previously associated with lung function (specifically
change in FVC, FEV1) in article "Epigenome-wide association study of lung
function level and its change", [PMID 31073081](https://erj.ersjournals.com/content/erj/early/2019/03/27/13993003.00457-2019.full.pdf). 

[Article](https://www.sciencedaily.com/releases/2013/10/131025091938.htm) about
an ASHG abstract in 2013 from Brigham and Women's Hospital: SNPs in XKR4 gene
associated with BDRPRED (change as % of predicted FEV1) & BDRBASE (change as %
of baseline FEV1).

### **Plot Model P-Values**

```{r, warning = F}
plot_sig(win_chr8)
```

### **AF**

*TODO: need to update automation for this section.*

## **maxFEV1/FVC Chr10**

### **LAAA Results**

**All:**

```{r}
# Get LAAA results & format
win_chr10_sig <- get_sig(win_chr10, nom_sig_thresh)

win_chr10_sig <- win_chr10_sig %>%
  dplyr::rowwise() %>%
  dplyr::mutate(chr = 10) %>%
  dplyr::relocate(anova_p, max_abs_beta) %>%
  dplyr::arrange(anova_p)

win_chr10_sig %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)
# datatable(win_chr10_sig)

```


**Summarized to keep only the first of those in complete LD:**

For now, doing this by filtering to keep only one of multiple rows with
identical alt_frq, anova_p, and max_abs_beta.

```{r}
win_chr10_sig_dedup <- win_chr10_sig %>%
  distinct(alt_frq, anova_p, max_abs_beta, .keep_all = TRUE)

win_chr10_sig_dedup %>%
  dplyr::select(
    position, anova_p, max_abs_beta, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)
```

Before MAF filter was correctly applied, we had 11 deduplicated significant
results, now we have 7.

For locus Zoom plot, want one version for each of the "peaks". To do this,
need to narrow down the above to keep only top SNP from each set of physically
close SNPs.

```{r}
# Get ranges of points within 30,000bp of each other
win_chr10_sig_ranges <-
  IRanges(start = win_chr10_sig$position, end = win_chr10_sig$position)
win_chr10_sig_merged_ranges <-
  IRanges::reduce(win_chr10_sig_ranges, min.gapwidth = 30000)

# Add range group to original df
win_chr10_sig$group <-
  subjectHits(findOverlaps(win_chr10_sig_ranges, win_chr10_sig_merged_ranges))

# Now, modify df to keep only most significant pval in each group

# Group by the merged ranges and keep the row with the lowest p-value in each group
win_chr10_sig_peaks <- win_chr10_sig %>%
  dplyr::group_by(group) %>%
  dplyr::slice(which.min(anova_p)) %>%
  ungroup() %>%
  arrange(anova_p)

head(win_chr10_sig_peaks)
```

### **Locus Zoom Plot**

Only plot most significant peak SNP:

```{r}
# Calculate LD
chr10_ld <- Reduce(full_join, lapply(win_chr10_sig_peaks$position, function(pos) {
  return(calc_peak_ld(pos, win_chr10_sig, df_chr10))
}))

# Plot LD on locus zoom plot
win_chr10_ld <- as.data.frame(inner_join(
  win_chr10 %>% dplyr::mutate(chr = 10),
  chr10_ld,
  by = c("position" = "pos")))  # should keep all entries

chr10_peak <- win_chr10_sig_peaks$position[which.min(win_chr10_sig_peaks$anova_p)]


# Plot locus zoom
plot_chr10_lz <- plot_locuszoom(win_chr10_ld, "10", chr10_peak)
plot_chr10_lz

```


### **Plot with Phenotype**

One for each SNP crossing our significance threshold, arranged from most to least significant.

```{r out.width = '50%', fig.show = 'hold'}
# Make plots
# Using walk to suppress ## 1, 2, output during lapply
# walk(win_chr10_sig_peaks$position, ~ {
#   print(plot_split(pheno_mff, df_chr10, win_chr10_sig, .x, "maxFEV1_FVC"))
# })

```

### **Draft Plot for Publication**

*TODO: to be added if decide to include any/all of the above!*

### **BiomaRt Lookup**

```{r}
win_chr10_sig_dedup$chr = 10

result <- do.call(rbind, lapply(1:nrow(win_chr10_sig_dedup), function(i) {
  result <- getBM(
    attributes = c('ensembl_gene_id', 'hgnc_symbol', 'chromosome_name', 'start_position', 'end_position'),
    filters = c('chromosome_name','start','end'),
    values = list(win_chr10_sig_dedup$chr[i], win_chr10_sig_dedup$position[i], win_chr10_sig_dedup$position[i]),
    mart = mart)
}))

head(result)

```


+ PCDH15 is in the cadherin superfamily
+ JACI association between different SNP in PCDH15 for FEV1
  [PMID 23541324](https://www.jacionline.org/article/S0091-6749(13)00261-3/fulltext)
+ PCDH15 listed on dbGAP gene-trait association with asthma
[see here](https://maayanlab.cloud/Harmonizome/gene_set/Asthma/dbGAP+Gene-Trait+Associations)

+ ENSG00000236958 appears to be a long non-coding RNA.


### **Plot Model P-Values**

```{r}
plot_sig(win_chr10)
```

### **AF**

*TODO: need to update automation for this section.*


## **MaxFVC Chr17**

### **LAAA Results**

Looking at all results meeting at least nominal significance: `r nom_sig_thresh`

```{r}
# Get sig LAAA results & format
win_chr17_sig <- get_sig(win_chr17, nom_sig_thresh)

win_chr17_sig <- win_chr17_sig %>%
  dplyr::rowwise() %>%
  dplyr::mutate(chr = 17) %>%
  dplyr::relocate(anova_p, max_abs_beta) %>%
  dplyr::arrange(anova_p)

win_chr17_sig %>%
  dplyr::select(
    position, anova_p, max_abs_beta, position, allele_dose_beta, allele_dose_p,
    afr_dose_beta, afr_dose_p, allele_afr_dose_beta, allele_afr_dose_p)
# datatable(win_chr17_sig)

```

This result is new - chr17 maxFVC before MAF filter didn't have any results.

For locus Zoom plot, want one version for each of the "peaks". To do this,
need to narrow down the above to keep only top SNP from each set of physically
close SNPs.

```{r}
# Get ranges of points within 30,000bp of each other
win_chr17_sig_ranges <-
  IRanges(start = win_chr17_sig$position, end = win_chr17_sig$position)
win_chr17_sig_merged_ranges <-
  IRanges::reduce(win_chr17_sig_ranges, min.gapwidth = 30000)

# Add range group to original df
win_chr17_sig$group <-
  subjectHits(findOverlaps(win_chr17_sig_ranges, win_chr17_sig_merged_ranges))

# Now, modify df to keep only most significant pval in each group

# Group by the merged ranges and keep the row with the lowest p-value in each group
win_chr17_sig_peaks <- win_chr17_sig %>%
  dplyr::group_by(group) %>%
  dplyr::slice(which.min(anova_p)) %>%
  ungroup() %>%
  arrange(anova_p)

head(win_chr17_sig_peaks)
```

### **Locus Zoom Plot**

Only plot most significant peak SNP:

```{r}
# Calculate LD
chr17_ld <- Reduce(full_join, lapply(win_chr17_sig_peaks$position, function(pos) {
  return(calc_peak_ld(pos, win_chr17_sig, df_chr17))
}))

# Plot LD on locus zoom plot
win_chr17_ld <- as.data.frame(inner_join(
  win_chr17 %>% dplyr::mutate(chr = 17),
  chr17_ld,
  by = c("position" = "pos")))  # should keep all entries

chr17_peak <- win_chr17_sig_peaks$position[which.min(win_chr17_sig_peaks$anova_p)]

# Plot locus zoom
plot_chr17_lz <- plot_locuszoom(win_chr17_ld, "17", chr17_peak)
plot_chr17_lz

```


### **Plot with Phenotype**

One for each SNP crossing our significance threshold, arranged from most to least significant.

```{r out.width = '50%', fig.show = 'hold'}
# Make plots
# Using walk to suppress ## 1, 2, output during lapply
# walk(win_chr17_sig_peaks$position, ~ {
#   print(plot_split(pheno_mfvc, df_chr17, win_chr17_sig, .x, "maxFVC"))
# })

```

### **Plot Model P-Values**

```{r}
plot_sig(win_chr17)
```


## **Compare Allele Estimate**

Want to compare full LAAA model to model run with only allele term and
covariates. Looking at this for all nominally significant results from both
regions combined, colored by chr/phenotype.

*TO NOTE: this is from older version of report, only looks at chr8 & chr10.*

```{r}
win_chr8_sig$pheno <- "maxFVC"
win_chr10_sig$pheno <- "maxFEV1_FVC"
sig_comb <- full_join(win_chr8_sig, win_chr10_sig)

ggplot(sig_comb, aes(x = allele_dose_beta, y = unadj_allele_dose_beta,
                     color = pheno)) +
  geom_point(alpha = 0.5) +
  coord_fixed() +
  xlim(c(-4,4)) +
  ylim(c(-4,4))
```

Also look at as table:

```{r}
sig_comb %>%
  arrange(pheno, position) %>%
  dplyr::select(pheno, position, unadj_allele_dose_beta, unadj_allele_dose_p,
                allele_dose_beta, allele_dose_p)
```




# **Plot for Publication**

### **Figure 4**

```{r}
# Switch maxFVC column from L to mL
pheno_mfvc <- pheno_mfvc %>%
  dplyr::mutate(maxFVC_L = maxFVC,
                maxFVC = maxFVC * 1000)

sig_pos <- win_chr8_sig$position[which.min(win_chr8_sig$anova_p)]

colors = c(color_lfvc, color_bfvc, color_maxfvc)
plot_chr8_pheno <- plot_split(
  pheno_mfvc, df_chr8, win_chr8_sig, sig_pos, "maxFVC", colors)

# plot_chr8_lz
plot_chr8_pheno + theme(plot.title = element_blank())

p_comb <- ggarrange(
  plot_chr8_lz,
  plot_chr8_pheno + theme(plot.title = element_blank()),
  labels = c("A", "B"), ncol=2, nrow=1)

ggsave(
  "dissemination/SARP_ancestry_paper_draft/figures/fig4_laaa_maxfvc_top_hit.png",
  p_comb, units = c("in"), width = 10, height = 5)

```

### **Supplemental Figure 3**

```{r}

### Chr10 MaxFEV1/FVC
sig_pos <- win_chr10_sig$position[which.min(win_chr10_sig$anova_p)]

colors = c(color_lratio, color_bratio, color_maxratio)
plot_chr10_pheno <- plot_split(
  pheno_mff, df_chr10, win_chr10_sig, sig_pos, "maxFEV1_FVC", colors)

# plot_chr10_lz
plot_chr10_pheno + theme(plot.title = element_blank())

### Chr17 MaxFVC
sig_pos <- win_chr17_sig$position[which.min(win_chr17_sig$anova_p)]

colors = c(color_lfvc, color_bfvc, color_maxfvc)
plot_chr17_pheno <- plot_split(
  pheno_mfvc, df_chr17, win_chr17_sig, sig_pos, "maxFVC", colors)

# plot_chr8_lz
plot_chr17_pheno + theme(plot.title = element_blank())

### Combine
p_comb <- ggarrange(
  plot_chr10_lz,
  plot_chr10_pheno + theme(plot.title = element_blank()),
  plot_chr17_lz,
  plot_chr17_pheno + theme(plot.title = element_blank()),
  labels = c("A", "B", "C", "D"), ncol=2, nrow=2)

ggsave(
  "dissemination/SARP_ancestry_paper_draft/figures/fig_s3_laaa_maxfvc_maxratio_hits.png",
  p_comb, units = c("in"), width = 10, height = 10)

```



# **Write Out**

Specifically writing out all SNPs that passed our significance threshold for
Abhishek to use in replication analysis.

Also adding biomaRt annotation of RSID here first.

## **Chr8**

```{r}
# Use biomaRt to annotate RSID
win_chr8_sig <- win_chr8_sig %>%
  dplyr::mutate(chr_pos_pos = paste0(chr, ":", position, ":", position),
                chr_pos_ref_alt = paste0("chr", chr, ":", position, ":", ref, ":", alt))

snp_mart <- useEnsembl(
  biomart="ENSEMBL_MART_SNP", dataset="hsapiens_snp", mirror = "useast")

win_chr8_sig_result <- getBM(
  attributes = c('refsnp_id', 'allele', 'chrom_start'),
  filters = 'chromosomal_region',
  values = win_chr8_sig$chr_pos_pos,
  mart = snp_mart)

win_chr8_sig_mod <- left_join(
  win_chr8_sig, win_chr8_sig_result, by = c("position" = "chrom_start")) %>%
  dplyr::mutate(ref_alt = paste0(ref, "/", alt))
```

For biomaRt RSID annotation, checking those RSIDs where position was identified
more than once:

```{r}
chr8_dup_pos <- win_chr8_sig_mod %>%
  count(position) %>%
  dplyr::filter(n > 1) %>%
  pull(position) # only 1

win_chr8_sig_mod %>% dplyr::filter(position %in% chr8_dup_pos) %>%
  dplyr::select(position, chr_pos_ref_alt, ref_alt, refsnp_id, allele)

# Keep only one match, add pheno
win_chr8_sig_mod_exp <- win_chr8_sig_mod %>%
  dplyr::filter(!(position == chr8_dup_pos & refsnp_id != "rs7820046")) %>%
  dplyr::mutate(pheno = "maxFVC")
# n_distinct(win_chr8_sig_mod$position)  # 24 / 24
```

Based on gnomAd and ref/alt alleles, RSID rs7820046 should be the correct match
for chr8:55567655:T:C.

## **Chr10**

```{r}
# Use biomaRt to annotate RSID
win_chr10_sig <- win_chr10_sig %>%
  dplyr::mutate(chr_pos_pos = paste0(chr, ":", position, ":", position),
                chr_pos_ref_alt = paste0("chr", chr, ":", position, ":", ref, ":", alt))

win_chr10_sig_result <- getBM(
  attributes = c('refsnp_id', 'allele', 'chrom_start'),
  filters = 'chromosomal_region',
  values = win_chr10_sig$chr_pos_pos,
  mart = snp_mart)

win_chr10_sig_mod <- left_join(
  win_chr10_sig, win_chr10_sig_result, by = c("position" = "chrom_start")) %>%
  dplyr::mutate(ref_alt = paste0(ref, "/", alt))
```

For biomaRt RSID annotation, checking those RSIDs where position was identified
more than once:

```{r}
chr10_dup_pos <- win_chr10_sig_mod %>%
  count(position) %>%
  dplyr::filter(n > 1) %>%
  pull(position) # 3

win_chr10_sig_mod %>% dplyr::filter(position %in% chr10_dup_pos) %>%
  dplyr::select(position, chr_pos_ref_alt, ref_alt, refsnp_id, allele)

# Keep only one match, add pheno
win_chr10_sig_mod_exp <- win_chr10_sig_mod %>%
  dplyr::filter(
    !(position == "55137025" & refsnp_id != "rs7905410"),
    !(position == "55137026" & refsnp_id != "rs7916351"),
    !(position == "56517980" & refsnp_id != "rs939657366")) %>%
  dplyr::mutate(pheno = "maxFEV1_FVC")
# n_distinct(win_chr10_sig_mod$position)  # 15 / 15
```

Based on gnomAd (or dbSNP) and ref/alt alleles:

+ RSID rs7905410 should be the correct match for chr10:55137025:T:C
+ RSID rs7916351 should be the correct match for chr10:55137026:G:A
+ RSID rs939657366 should be the correct match for chr10:56517980:T:A

## **Chr17**

```{r}
# Use biomaRt to annotate RSID
win_chr17_sig <- win_chr17_sig %>%
  dplyr::mutate(chr_pos_pos = paste0(chr, ":", position, ":", position),
                chr_pos_ref_alt = paste0("chr", chr, ":", position, ":", ref, ":", alt))

# BiomaRt failing for this SNP - try manual lookup on gnomAd
# win_chr17_sig_result <- getBM(
#   attributes = c('refsnp_id', 'allele', 'chrom_start'),
#   filters = 'chromosomal_region',
#   values = win_chr17_sig$chr_pos_pos,
#   mart = snp_mart)
# win_chr17_sig$chr_pos_ref_alt  # "chr17:79396701:A:G"
# rs907901

win_chr17_sig_mod <- win_chr17_sig %>%
  dplyr::mutate(refsnp_id = "rs907901",
                pheno = "maxFVC")
```

## **Write Out All Chr**

*TO NOTE: re-loaded output significant results and added max_abs_beta here*
*because at time biomaRt was stalling so couldn't re-run full report.*

```{r}
merge_temp <- full_join(
  win_chr8_sig_mod_exp,
  win_chr10_sig_mod_exp)  # should be 15 + 24 = 39
merge <- full_join(
  merge_temp,
  win_chr17_sig_mod)  # should be 39 + 1 = 40

# Modify all for export
# Group here was used to group plotting windows, so removing when write out
merge_exp <- merge %>%
  dplyr::select(-chr, -group, -chr_pos_pos, -allele, -ref_alt, -max_abs_beta) %>%
  dplyr::relocate(anova_p, .after = "n") %>%
  dplyr::relocate(pheno, chr_pos_ref_alt, refsnp_id)

# merge_exp <- read_tsv("data/laaa_wgs_new_2/output/laaa_significant_results.txt")

merge_exp$max_abs_beta <- apply(
    merge_exp[, c("allele_dose_beta", "afr_dose_beta", "allele_afr_dose_beta")],
    1, function(row) {
      cols <- c("allele_dose_beta", "afr_dose_beta", "allele_afr_dose_beta")
      cols[which.max(abs(row))]
  })

merge_exp <- merge_exp %>%
  dplyr::relocate(max_abs_beta, .after = pheno)

# Write out
write_tsv(
  merge_exp,
  "data/laaa_wgs_new_2/output/laaa_significant_results.txt")
```


# **Session Info**

```{r session_info}
sessionInfo()
```
