---
title: "Local ancestry inference QC"
author: "SDS, modified from MPB"
date: "Last compiled on `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: kable
    code_folding: hide
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../reports") })
---


```{r setup, include = F}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(
  root.dir = paste0(
    "/Users/slacksa/OneDrive - The University of Colorado Denver/",
    "Collabs/ortega"))

library(tidyverse)
library(conflicted)

```


```{r, message = F}
# TODO: add paths
# Path to "merged_gwide.txt" file made by pipeline
gwide <- read.delim("data/rfmix/output/merged_gwide.txt")
# Path to "local_ancestry_segments.txt" file made by pipeline
local.anc <- read.delim("data/rfmix/output/local_ancestry_segments.txt")
# Path to RFMix output dir
rfmix.dir <- "data/rfmix/output"

```


## Genome-wide ancestry differences

Below is a summary and histogram of the differences between genome-wide ancestry
estimated using ADMIXTURE, and genome-wide ancestry estimated using RFMix, per
individual.

```{r, echo=FALSE}
delta <- gwide$RFMIX_GW_AFR - gwide$ADMIXTURE_GW_AFR
summary(delta)
hist(delta, breaks=50, main="Genome-wide ancestry estimate differences")
```

```{r}
ggplot(gwide, aes(RFMIX_GW_AFR, ADMIXTURE_GW_AFR)) +
  geom_point() +
  xlim(c(0,1)) +
  ylim(c(0,1))
```


## Local ancestry deviation distribution

The deviation between mean local African ancestry and mean genome-wide African
ancestry was  calculated for each local ancestry segment. The distribution of
these deviations look as expected.

```{r, echo=FALSE}
mean.afr.anc <- mean(gwide$RFMIX_GW_AFR)
test <- mean(gwide$ADMIXTURE_GW_AFR)
dev <- local.anc$prop.afr - mean.afr.anc
z <- dev/sd(dev)
hist(z, breaks=1000, main="Histogram of local ancestry deviations",
     xlim = c(-4, 3))
```

## Local ancestry across the genome

Below is a plot of the mean local ancestry across the genome.

```{r, echo=FALSE, fig.height=12, fig.width=9}
draw.frame <- data.frame()
for (chr in 1:22) {
  chr.anc <- local.anc[local.anc$chr.nr == chr,] 
  chr.anc$end.snp.nr <- cumsum(chr.anc$nr.snps)
  chr.anc$start.snp.nr <- c(1, chr.anc$end.snp.nr[-length(chr.anc$end.snp.nr)] + 1)
  snps <- read.table(paste0(rfmix.dir, "/chr", chr, "/chr", chr, "_local_ancestry_snps.txt"))[,1]
  chr.anc$begin.pos <- snps[chr.anc$start.snp.nr]
  chr.anc$end.pos <- snps[chr.anc$end.snp.nr]
  chr.anc <- chr.anc[order(chr.anc$seg.nr),]
  chr.anc$pos <- rowMeans(chr.anc[,c(dim(chr.anc)[2]-1,dim(chr.anc)[2])])
  draw.frame <- rbind(draw.frame,
                      data.frame(Chromosome=paste("Chromosome", rep(chr, dim(chr.anc)[1])),
                                 Position=chr.anc$pos/10000000,
                                 Ancestry=chr.anc$prop.afr.anc))

}
draw.frame$Chromosome <- factor(draw.frame$Chromosome, 
                                levels=c(
                                  "Chromosome 1", "Chromosome 2", "Chromosome 3", "Chromosome 4",
                                  "Chromosome 5", "Chromosome 6", "Chromosome 7", "Chromosome 8",
                                  "Chromosome 9","Chromosome 10","Chromosome 11","Chromosome 12",
                                  "Chromosome 13","Chromosome 14","Chromosome 15","Chromosome 16",
                                  "Chromosome 17","Chromosome 18","Chromosome 19","Chromosome 20",
                                  "Chromosome 21","Chromosome 22"))
library(ggplot2)
ggplot(draw.frame, aes(Position, Ancestry)) +  
  facet_wrap( ~ Chromosome, nrow=6, ncol=4) + 
  geom_line(size=0.2) + 
  labs(x = expression(paste("Base pair position ", (10^7)))) + 
  labs(y = "Proportion African ancestry") +
  theme_bw() +
  theme(legend.position="top") + 
  scale_color_brewer(palette="Set1") +
  geom_hline(aes(yintercept=mean.afr.anc), size=0.0725)
```

